### 2024-04-15 outputs/paper/paper.qmd
set a connection

---
title: "Better Talent and Rule Changes Have Resulted in More Goals in the NHL"
subtitle: "An Analysis of the Increase in NHL Goal Scoring From 2011 to 2023"
author: 
  - Jason Ngo
thanks: "Code and data are available at: https://github.com/thejasonminh/nhlpoints"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "Of the four major sports in the USA, hockey is the least popular. In an effort to grow popularity, the NHL (National Hockey League) strove to increase goal scoring. Using data from the NHL, this paper will analyze average leaguewide goal scoring per game from the 2011-2012 to 2022-2023 seaons. A linear regression model will also be created to predict goal scoring averages in the upcoming season. Results show that the NHL was succesful as the average goals per game increased over a 10 year period. These findings suggest that increased player talent and rule changes are primary factors."
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)

league_data <- read.csv(here::here("outputs/data/league_data.csv"))
master_dmen <- read.csv(here::here("outputs/data/master_dmen.csv"))
master_forwards <- read.csv(here::here("outputs/data/master_forwards.csv"))
goals_model <- read_rds(here::here("outputs/models/goals_model.rds"))
```

# Introduction
Ice Hockey, Baseball, Football, and Basketball have become so popular in the United States, they've become branded as "The Big Four." The top leagues that operate in each sport: The NHL, MLB, NFL, and NBA are all centralized in the USA. Many of the top players that have played, or currently play, in these leagues are American, and their popularity only serves to inspire more youth to follow in their footsteps. There are no other sports that can threaten the stranglehold that these four sports have on the USA.

However, hockey has fallen behind the other three. Viewership, revenue and general popularity all pale in comparison to the other three sports [@citestatista]. As a result; player salaries are lower in comparison to other sports and there is less youth talent in the pipelines. Two of the biggest factors for this decline are: High price of entry and low scoring. @citegear details the high upfront price of equipment and the amount of equipment required to play ice hockey. Because of this high upfront price, many kids elect not to choose hockey as their sport and instead elect for cheaper options such as basketball. Low scoring was a result of goalies having too much freedom, and teams creating bulletproof defensive tactics that involved clogging the middle of the ice and using their goalie to stifle momentum.

As a response to the declining popularity, the NHL decided to implement rule changes and goalie equipment resizing to facilitate the increase of goal scoring. By increasing goal scoring, games would be more exciting to watch which in turn, would lead to increased viewership. This paper will analyze the effectiveness of these changes by drawing data from the NHL to analyze the estimand: Average leaguewide goals scored per game from the 2011-2012 season to the current (2022-2023) season. In addition, the paper will draw on a linear regression model to predict average goal scoring for the future season.

The findings show an increase of average goals scored per game over a 12 season period, with this current season exhibiting the highest goals scored per game during this period. The rule changes have given players more freedom to play and goalie equipment resizing has made goalies more susceptible to getting scored on. In addition, player talent has greatly increased over the years, with both forwards and defensemen exhibiting drastic offensive growth.

In @sec-data of the paper, the source of the data as well as the three primary datasets are discussed. Strengths and weaknesses, methods of collection and terminologies will be explored. @sec-model of the paper will introduce a linear model, its justifications, and how it was constructed. In addition, the model will predict goal scoring averages for next season. @sec-results will discuss the results, touching on the trends and events that occurred over a 12-season period. @sec-discussion of the paper will touch on the factors that contributed to increased goal scoring as well as weaknesses and next steps.

# Data {#sec-data}
There are three primary datasets used for this paper. The first dataset, @sec-leagueavg, contains league averages from NHL seasons 2011-12 to 2022-23. The next dataset, @sec-fwds, contains the average points-per-game for every forward from the 2011-12 to 2022-23 seasons with a recorded average points-per-game pace of 1.00 or above. The last dataset, @sec-dmen, contains the average points-per-game for every defenseman from seasons 2011-12 to 2022-23 with a recorded average points-per-game pace of 0.60 or above.

The analysis of these three datasets will be carried out using the statistical programming language `R` [@citeR], using the `tidyverse`[@citetidy], `here`[@citehere], and `readxl`[@citexl] packages. The figures and tables in the paper are generated using the, respectively, `ggplot2`[@citegg] and `knitr`[@citeknitr] packages.

## League Averages Dataset {#sec-leagueavg}

The league averages dataset, represented below in @tbl-1, is created through averaging each variable across every team in the league for the corresponding season. Seasons 2011-12 to 2016-17 has averages for 30 teams, seasons 2017-18 to 2020-21 has averages for 31 teams, and seasons 2021-22 and 2022-23 has averages for 32 teams. Each team is responsible for tracking their own statistics, which is then conglomerated and represented by both the NHL and third party websites. Statistics tracking is done during real time, while a game is in progress. Internal models are used to calculate percentage, average and advanced statistics.

With every team using their own methods for tracking statistics, measurement bias comes into play. Because the NHL does not standardize methods for tracking statistics, some teams will skew statistics such that their team is represented better.

```{r}
#| label: tbl-1
#| tbl-cap: The dataset containing averages from NHL seasons 2011-2012 to 2022-2023.
#| echo: false
#| warning: false

kable(league_data, padding = 5)
```

Almost all of @tbl-1 is taken from @citehref, with the exception of the `Sh_Pct` column, which was pulled from @citenhl and merged with the larger dataset. The reasoning why I didn't use the NHL website for the majority of this dataset was that the NHL website did not support automatic calculation of league averages. There are 8 columns total in this league averages dataset:

1. The `Season` column, which denotes the NHL season each average represents. 
2. The `Goals_PG` column, which denotes the average goals per game for the respective season. 
3. The `PP_PG` column, which denotes the average power plays, situations where one team has a man-advantage due to a penalty, per game for the respective season. 
4. The `PP_Pct` column, which shows the percentage of power plays that were successful, or resulted in a goal, for the respective season. 
5. The `SA` column, which details the leaguewide average shots a team faced per game, for the respective season.
6. The `SV_Pct` column, which denotes the leaguewide average save percentage, which represents the percentage of total saves made, per season.
7. The `GA_Avg` column, which represents the leaguewide average goals against per season.
8. The `Sh_Pct` columnm which represents the leaguewide average shooting percentage. Shooting percentage represents the percentage of shots that are goals.

The `Season` column was necessary as a timeframe is needed to anchor this analysis. The `Goals_PG` was chosen as it is the estimand of this paper. The remaining 6 columns were chosen as they all correlate to goal scoring, or situations that incentivizes goal scoring. For example, having a high shooting percentage correlates to having a higher goals per game, which would decrease the save percentage and increase the goals against average. The powerplay columns represent the opportunities teams have to score on the man advantage and the capability for teams to score on that advantage.

## Forwards Dataset {#sec-fwds}

The forwards dataset, represented below in @tbl-2, represents the count of every forward from the 2011-12 to 2022-23 seasons that finished the season with, at minimum, a 1 point-per-game (PPG) pace. A point represents either a goal or an assist. The reason why I chose to filter the forwards by the 1 PPG pace is because there are few players that finish the season at a 1 PPG pace, which makes those players that are able to do so offensively gifted.

The calculation of average points-per-game is quite simple: Simply take a player's total points and ratio it against the amount of games they've played. It is very simple to track a player's points since every game log has every goal with their respective assists recorded. This is true for both forwards and defensemen.

```{r}
#| label: tbl-2
#| tbl-cap: The count of forwards that finish at a 1 PPG pace or above per season.
#| echo: false
#| warning: false

kable(master_forwards |> count(Season) |> rename("PPG Players" = "n"), padding = 5)
```

All the data about forwards was taken from [@citenhl]. Since we are only concerned with the average offensive output of forwards, I chose to only analyze the points-per-game pace of forwards.

## Defensemen Dataset {#sec-dmen}

The defensemen dataset, represented below in @tbl-3, represents the count of every defensemen from the 2011-12 to 2022-23 seasons that finished the season with, at minimum, a 0.6 point-per-game pace. Rather than filtering by the 1 PPG pace like I did for forwards, I chose to filter defensemen by a 0.6 PPG pace. The reasoning for doing so is because defensemen simply do not contribute offensively as much as forwards do. A defensemen will rarely match a forward's offensive output because offense isn't the primary objective of a defensemen. However, talented defensemen that are table to contribute offensively usually do so at a 0.6 point-per-game pace or above.

```{r}
#| label: tbl-3
#| tbl-cap: The count of defensemen that finish at a 0.6 PPG pace or above per season.
#| echo: false
#| warning: false

kable(master_dmen |> count(Season) |> rename("PPG Players" = "n"), padding = 5)
```

All the data regarding defensemen was also taken from [@citenhl]. Similar to the forwards, I am only concerned with the offensive output of defensemen so I chose to only analyze the point-per-game pace of defensemen.

# Model {#sec-model}

The goal of my modelling strategy is twofold. Firstly, I want to analyze to what effect power play opportunities, power play effectiveness, save percentage, goals against average and shooting percentages all have on average goals scored per game. Secondly, I want to use these results to predict the average goals scored per game for the next season.

In the section below, the Bayesian analysis model used to investigate the multiple linear regression model of average goals scored per game is detailed.

## Model set-up

Define $y_i$ as the leaguewide average goals scored per game. Then $\beta_1$ is the leaguewide average power plays per game, $\gamma_1$ is leaguewide powerplay percentage, $\theta_1$ is leaguewide save percentage, $\phi_1$ is the leaguewide goals against average, and $\kappa_1$ is the leaguewide average shooting percentage.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \theta_i + \phi_i + \kappa_i + \beta_i * \gamma_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\theta &\sim \mbox{Normal}(0, 2.5) \\
\phi &\sim \mbox{Normal}(0, 2.5) \\
\kappa &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in `R` [@citeR] using the `rstanarm` package of [@citestan]. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between average goals scored per game and power play opportunities, power play percentage, goals against average, and shooting percentage. The higher those four variables are, the more goals there should be. In particular, there should also be a negative relationship between goals scored and save percentage. With a lower save percentage, there are less saves that are occuring which means there should be more goals. Additionally, the two power play variables are interacted with eachother, since they are directly related to one another. There cannot be a power play percentage without a power play opportunity.

## Model Prediction
Taking the differences in averages from the 2021-2022 season and the 2022-2023 season, I will create a test dataset for the 2023-2024 season by adding those differences onto the averages from the 2022-2023 season and use the model with this test dataset to predict the average goals scored per game for the 2023-2024 season. @tbl-4 is the resulting mock 2023-2024 season averages data.
```{r}
#| echo: false
#| warning: false

test_league <- tibble(
  Season = c(
    "2023-24"
    ),
  Goals_PG = c(
    3.22
  ),
  PP_PG = c(
    3.25
  ),
  PP_Pct = c(
    22.01
  ),
  SA = c(
    30.8
  ),
  SV_Pct = c(
    0.901
  ),
  GA_Avg = c(
    3.02
  ),
  Sh_Pct = c(
    10.33
  )
)

prediction <- predict(goals_model, newdata = test_league)
```

```{r}
#| label: tbl-4
#| tbl-cap: A mock league averages dataset for the 2023-2024 NHL season.
#| echo: false
#| warning: false

kable(test_league, padding = 5)
```

Using the mock data alongside the model, the predicted average goals per game for the 2023-2024 season is **3.21**.

\newpage

# Results {#sec-results}

## Goals, Saves and Shooting

```{r}
#| label: fig-1
#| fig-cap: An increase of average goals scored per game over 12 seasons.
#| echo: false
#| warning: false

library(tidyverse)

league_data |> ggplot(aes(x = Season, y = Goals_PG)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Leaguewide Average Goals Scored Per Game over 12 Seasons",
    x = "Season",
    y = "Average Goals P/G"
  ) 
```

@fig-1 shows an upward trajectory of goal scoring per season. The horizontal axis denotes the season, and the vertical axis denotes the average goals scored per game. There are two things of note: First, there was a big spike in goal scoring starting in the 2017-18 season, with an increase of 0.2 from the 2016-17 season. with the subsequent seasons building on that jump. The second thing of note however, is the slight dip in scoring during the 2020-21 season where the average fell to 2.94 from the 3.02 the season before.

In regards to the spike of scoring in the 2017-18 season, the NHL had implemented a few rule changes that helped facilitate scoring. Additionally, goalie equipment had shrunken by the time the season started, so goalies were scored on more.

With respect to the slight dip in scoring in the 2020-21 season, that season took place during the height of the COVID-19 pandemic and due to health regulations, many players had to miss extended periods of time to self-quarantine after a positive test. The loss of skilled players for periods of time and the erosion of team chemistry led to the slight decrease in goals scored that season.

\newpage

```{r}
#| label: fig-2
#| fig-cap: Implications on increased goal scoring on save percentage, goals against average and shooting percentage.
#| fig-subcap: ["Save Percentage", "Goals Against Average", "Shooting Percentage"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(tidyverse)

league_data |> ggplot(aes(x = Season, y = SV_Pct)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Leaguewide Save Percentage over 12 seasons",
    x = "Season",
    y = "Average Save %"
  )

league_data |> ggplot(aes(x = Season, y = GA_Avg)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Leaguewide Goals Against Average over 12 seasons",
    x = "Season",
    y = "GAA"
  )

league_data |> ggplot(aes(x = Season, y = Sh_Pct)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Leaguewide Average Shooting Percentage over 12 seasons",
    x = "Season",
    y = "Average Sh %"
  ) 
```

@fig-2 shows the implications of increased goal scoring in average save percentage, goals against average and average shooting percentage. A better shooting percentage means more goals are scored. More goals scored increases the goals against average, and more goals scored means less saves are made.

\newpage

## More Talent, Better Power Plays

```{r}
#| label: fig-3
#| fig-cap: The rise of offensively gifted forwards and defensemen over 12 seasons.
#| fig-subcap: ["Forwards", "Defensemen", "Shots Per Game"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(tidyverse)

master_forwards |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Productive Forwards over 12 seasons",
    x = "Season",
    y = ">= 1 PPG Forwards"
  )

master_dmen |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Productive Defensemen over 12 seasons",
    x = "Season",
    y = ">= 0.6 PPG Defensemen"
  )

league_data |> ggplot(aes(x = Season, y = SA)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Avg. Shots Per Game over 12 seasons",
    x = "Season",
    y = "Shots Per Game"
  ) 
```

@fig-3 showcases the increase of offensively talented forwards and defensemen over 12 seasons, with an additional graph of leaguewide shots per game to contextualize the increase in talent. For offensively talented forwards, the horizontal axis denotes the season while the vertical axis denotes the amount of forwards that finished the season at a pace of 1 PPG or greater. Similarly, the offensively talented defensemen graph has the same horizontal axis, while the vertical axis denotes the amount of defensemen that finished the season at a pace of 0.6 PPG or greater. For shots per game, the horizontal axis once again denotes the season and the vertical axis denotes the average leaguewide shots per game.

Similar to the above section about goal scoring, the shift from the 2016-17 to the 2017-18 season brought an uptick in scoring alongside it. Regarding forwards, it is interesting to note that 2012-13 total of 20 players was never reached again until the 2017-18 season. The 2014-15, 2015-16, and 2016-17 seasons all saw a drop in productive forwards. This trend is not reflected in the defensemen graph. Aside from a slight drop in the 2016-17 season, the amount of offensively productive defensemen slowly increased over the years. This trend means that over the seasons, defensemen have gotten more involved with the offense. This could be a result of increased talent, or a shifting in tactics to allow defensemen to flourish offensively, or perhaps a mixture of both.

However, what makes this increase in productivity even more substantial is looking at the average shots per game over 12 seasons. Over the 12 season period, the lowest average shots per game was 29, and thee highest was 31.8. This means that teams are barely shooting more; the average shots per game over 12 seasons has remained somewhat consistent. Players are simply doing more with their shots on goal.

\newpage

```{r}
#| label: fig-4
#| fig-cap: The implications of better talent on power play effectiveness.
#| fig-subcap: ["Amount of Power Plays", "Effectiveness of Power Plays"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(tidyverse)

league_data |> ggplot(aes(x = Season, y = PP_PG)) + 
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Power Play Opportunities over 12 seasons",
    x = "Season",
    y = "Avg. PP Opportunities"
  )

league_data |> ggplot(aes(x = Season, y = PP_Pct)) + 
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Power Play Effectiveness over 12 seasons",
    x = "Season",
    y = "Avg. PP %"
  ) 
```

One consequence of increased offensive talent is a greater effectiveness of the power play. @fig-4 shows that, despite a decrease in overall power plays over a 12 season period, power play effectiveness has trended upwards. In essence, players are making more with less, and are doing so because of how much more skilled players are in comparison to players before.

# Discussion {#sec-discussion}

As evidenced by the results, goal scoring has steadily increased over the seasons. The NHL was successful in their goal of making the game more exciting, but how did they go about doing it? The two primary factors that led to this increase was an increase of mid-to-high tier player talent, as well as rule changes.

\newpage

## Good Players Becoming Great

While I talked about the increase of the amount of productive forwards and defensemen, here I will discuss the increase in the amount of good forwards and compare them to the amount of elite forwards. To be classified as an elite forward, usually they would need to finish the season at around a 1.4 PPG pace. Because of how hard of a feat it is to achieve a sustained 1.4 PPG pace, there are very few elite forwards in the NHL. @fig-5 shows the amount of forwards that finish in multiple PPG ranges.

```{r}
#| label: fig-5
#| fig-cap: Different Forward PPG brackets over 12 seasons.
#| fig-subcap: ["1.4 or above", "Between 1.2 and 1.4", "Between 1.0 and 1.2"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(tidyverse)

master_forwards |> filter(Points_PG >= 1.4) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Forwards above 1.4 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  )

master_forwards |> filter(Points_PG >= 1.2 & Points_PG < 1.4) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Forwards Between 1.2 and 1.4 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  )

master_forwards |> filter(Points_PG >= 1.0 & Points_PG < 1.2) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Forwards Between 1.0 and 1.2 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  ) 
```

As evidenced, the amount of elite players have remained consistent over the seasons, although the big spike in the 2021-22 season may be a sign that times are changing. The bigger story here though, is the amount of players that finish in the 1.2 and 1.4 range from 2017-18 beyond. There is a dramatic spike in these productive players from before 2017-18 to that season and beyond. The graphs here show that there while the ceiling of player talent has remained the same, the floor is steadily increasing.

\newpage

The same level of growth can be seen with defensemen as well. @fig-6 shows the amount of defensemen that finish in multiple PPG brackets.

```{r}
#| label: fig-6
#| fig-cap: Different Defenseman PPG brackets over 12 seasons.
#| fig-subcap: ["0.95 or above", "Between 0.75 and 0.95", "Between 0.6 and 0.75"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(tidyverse)

master_dmen |> filter(Points_PG >= 0.95) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Defensemen above 0.95 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  )

master_dmen |> filter(Points_PG >= 0.75 & Points_PG < 0.95) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Defensemen Between 0.75 and 0.95 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  )

master_dmen |> filter(Points_PG >= 0.6 & Points_PG < 0.75) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Defensemen Between 0.6 and 0.75 PPG over 12 seasons",
    x = "Season",
    y = "Player Count"
  ) 
```
There is a slightly different trajectory with defensemen. It appears that while the floor of offensive production has remained somewhat consistent over the seasons, the ceiling has only gotten higher. It appears that we may be in the middle of a renaissance of offensively gifted defensemen, with multiple defensemen over the last two seasons exhibiting a PPG of 0.95 or above. To contextualize, a great forward should finish at a 1.00 PPG, so to see multiple defensemen reach those numbers is unprecedented.

The floor of talent for forwards have increased, and the ceiling of talent for defensemen have increased, and both of these factors contributed to a spike in goal scoring.

\newpage

## Changes in 2017-18

While the findings have shown a general increase in goal scoring, the biggest spike in productivity lies between the 2016-17 season and the 2017-18 season. While blossoming player talent is certainly a factor here, the biggest factor in play are rule changes and goalie equipment changes.

### Rule Changes

@citerules details all of the rule changes effective for the beginning of the 2017-18 season. These changes include a power play consequence for a failing an off-side challenge (forcing an analyzed replay to see if a play was offside), denying time-outs for teams committing icing (throwing the puck down the entire length of the ice to delay the game and to change shifts), face-off placement changes and stricter enforcement on the stick-slashing around a player's hands. All of these changes, save for the face-off placement changes, had an effect on goal scoring. The off-side challenge rule change introduced more power plays, denying time-outs prevents tired players from resting up during icing infractions which led to worse defense, and the enforcement of slashing led to more offensive freedom and creativity for players skilled at working with the puck.

### Goalie Equipment Changes

@citegoalies shows the timeline of NHL goalie equipment changes. These changes included making goalie equipment and padding smaller to facilitate more scoring. While the changes were put into effect during the latter half of the 2016-17 season, teams became accustomed to the adjusted size and found more ways to score on these smaller goalies in the 2017-18 season. This is shown in the increase of goals against average and the drop in save percentage.

## Weaknesses

The primary weakness of my analysis is how surface level it is. I only focused on basic statistics such as per-game and per-season averages. A more richer and defined analysis would've been done if advanced statistics were used in this analysis. Analysis on these advanced stats would've been conducted on a team level and on an individual level to further investigate the causes for the increase in goal scoring.

Another weakness of my analysis is that it covers leaguewide averages per season. One problem with taking leaguewide averages is that in the NHL, there is very wide gap in skill between the teams at the top and teams at the bottom. Due to how drafting works in the NHL, teams at the bottom of the standings are encouraged to play worse in hopes of a better chance at drafting better players. In an analysis of leaguewide goal scoring, it feels wrong to include teams that are prioritizing losing over winning for a better future.

## Next Steps

An appropriate next step of this analysis is to compare the increase in goal scoring with viewership numbers, to see if the NHL was successful in drawing more eyes to the game. Additionally, to further analyze viewership rates, analysis on streaming options, media narratives and rivalry match-ups is recommended. Alongside the analysis on viewership, analysis on the amount of youth talent based on national viewership in not just America, but for European countries as well, is recommended to determine the outlook of the future of the NHL.

\newpage

# References

### 2024-04-15 scripts/00-simulate_data.R

@@ -1,0 +1,57 @@
+#### Preamble ####
+# Purpose: Simulates the primary datasets of interest
+# Author: Jason Ngo
+# Data: Apr 16 2023
+# Contact: jason_ngo@live.com
+# License: MIT
+# Pre-requisites: Some knowledge of NHL statistics terminology
+
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(2223)
+
+#### Simulate data ####
+## Season column correlates to the NHL season. Column XX represents season 20XX - 20YY where YY is (XX  + 1) ##
+## For example, column 16 represents season 2016 - 2017
+
+### Tibble for the league data simulation ###
+## Avg_Goals_Game has a min of 1, since realistically every team should average 1 goal a game atleast ##
+## Avg_Shots_Game has a min of 20 and max of 45, since those averages are hard to come by in a normal NHL season
+## Avg_PP_Percentage has a min of 5 and a max of 40, for the same reasons above
+## Avg_Sh_Percentage has a min of 5 and a max of 20, for the same reasons above
+nhlleaguesim <- tibble(
+  Season = c(
+    11:22
+  ),
+  Avg_Goals_Game = c(
+    runif(n = 12, min = 1, max = 5)
+  ),
+  Avg_Shots_Game = c(
+    runif(n = 12, min = 20, max = 45)
+  ),
+  Avg_PP_Percentage = c(
+    runif(n = 12, min = 5, max = 40)
+  ),
+  Avg_Sh_Percentage = c(
+    runif(n = 12, min = 5, max = 20)
+  ),
+  Avg_PP_Opportunities = c(
+    runif(n = 12, min = 1, max = 6)
+  )
+)
+
+### Tibble for player simulation data ###
+## Fwd_Pts_Game, since we are looking for point per game players, has a min of 1 and a max of 2 ##
+## Dmen_Pts_Game, since we are looking for productive defensemen, has a min of 0.5 and a max of 1.5 ##
+nhlplayersim <- tibble(
+  Season = c(
+    11:22
+  ),
+  Fwd_Pts_Game = c(
+    runif(n = 12, min = 1, max = 2)
+  ),
+  Dmen_Pts_Game = c(
+    runif(n = 12, min = 0.5, max = 1.5)
+  )
+)

### 2024-04-15 scripts/01-download_data.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Describes the acquisition process of the datasets used
+# Author: Jason Ngo
+# Data: Apr 16 2023
+# Contact: jason_ngo@live.com
+# License: MIT
+# Pre-requisites: A working internet connection
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Download data ####
+
+## League averages data downloaded from https://www.hockey-reference.com/leagues/stats.html#stats ##
+## Data on shooting is from nhl.com/stats/teams, from seasons 2011-2012 to 2022-2023 and viewing Shots By Type report ##
+## Data on forwards and dmen is from nhl.com/stats/skaters, from seasons 2011-2012 to 2022-2023 with 15 minimum games ## ##
+## forwards data was selected by filtering all players with 1.0 points per game or above from those seasons ##
+## defensemen data was selected by filtering all players with 0.6 points per game or above from those seasons ##
+
+# I acknowledge that the nhl api exists and is a usable R package, however.. #
+# ..isolating specific statistics by teams and players during a 10 year period using the API was too difficult #

### 2024-04-15 scripts/02-data_cleaning.R

@@ -1,0 +1,129 @@
+#### Preamble ####
+# Purpose: Isolates the columns of interest, sums the data and cleans up the names of columns.
+# Author: Jason Ngo
+# Data: Apr 16 2023
+# Contact: jason_ngo@live.com
+# License: MIT
+# Pre-requisites: Some knowledge of NHL statistics terminology
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(here)
+library(readxl)
+
+league_data <- read_excel(here::here("inputs/data/Summaryleague_1122.xlsx"))
+
+## The datasets from nhl.com come in pages, so I will load them and merge them immediately ##
+shooting_p1 <- read_excel(here::here("inputs/data/shooting_1122p1.xlsx"))
+shooting_p2 <- read_excel(here::here("inputs/data/shooting_1122p2.xlsx"))
+shooting_p3 <- read_excel(here::here("inputs/data/shooting_1122p3.xlsx"))
+shooting_p4 <- read_excel(here::here("inputs/data/shooting_1122p4.xlsx"))
+master_shooting <- rbind(shooting_p1, shooting_p2) |> rbind(shooting_p3) |> rbind(shooting_p4)
+
+forwards_p1 <- read_excel(here::here("inputs/data/forwards_1122p1.xlsx"))
+forwards_p2 <- read_excel(here::here("inputs/data/forwards_1122p2.xlsx"))
+forwards_p3 <- read_excel(here::here("inputs/data/forwards_1122p3.xlsx"))
+master_forwards <- rbind(forwards_p1, forwards_p2) |> rbind(forwards_p3)
+
+dmen_p1 <- read_excel(here::here("inputs/data/dmen_1122p1.xlsx"))
+dmen_p2 <- read_excel(here::here("inputs/data/dmen_1122p2.xlsx"))
+dmen_p3 <- read_excel(here::here("inputs/data/dmen_1122p3.xlsx"))
+master_dmen <- rbind(dmen_p1, dmen_p2) |> rbind(dmen_p3)
+
+#### Clean data ####
+
+### Fixing up the league_data dataset and renaming columns ###
+league_data[12,1] = "2011-12"
+league_data <- league_data |> rename(
+  "Goals_PG" = "G",
+  "PP_PG" = "PPO",
+  "PP_Pct" = "PP%",
+  "SA" = "SA",
+  "SV_Pct" = "SV%",
+  "GA_Avg" = "GAA"
+)
+
[…]

### 2024-04-15 scripts/03-test_data.R

@@ -1,0 +1,52 @@
+#### Preamble ####
+# Purpose: Tests the master dataframes against bound checks defined in the simulation script
+# Author: Jason Ngo
+# Data: Apr 16 2023
+# Contact: jason_ngo@live.com
+# License: MIT
+# Pre-requisites: Some knowledge of NHL statistics terminology
+
+#### Workspace setup ####
+library(tidyverse)
+library(here)
+
+league_data <- read.csv(here::here("outputs/data/league_data.csv"))
+master_dmen <- read.csv(here::here("outputs/data/master_dmen.csv"))
+master_forwards <- read.csv(here::here("outputs/data/master_forwards.csv"))
+
+#### Test data ####
+
+#### League dataset tests ####
+
+## Let's test if there are 12 columns, aka 12 seasons, in the league dataset ##
+league_data |> unique() |> count() == 12
+
+## Let's test if average goals is less than 5 and greater than 1. AKA is it a realistic value ##
+league_data$Goals_Per_Game < 5
+league_data$Goals_Per_Game > 1
+
+## Let's test if average shots per game is less than 45 and greater than 20. AKA is it a realistic value ##
+league_data$Shots_Against > 20
+league_data$Shots_Against < 45
+
+## Let's test if average PP percentage is less than 40 and greataer than 5. AKA is it a realistic value ##
+league_data$PowerPlay_Pct > 5
+league_data$PowerPlay_Pct < 40
+
+## Let's test if mean shooting percentage is less than 20 and greataer than 5. AKA is it a realistic value ##
+league_data$Mean_Sh_Pct > 5
+league_data$Mean_Sh_Pct < 20
+
+## Let's test if average powerplay opportunities/game is less than 6 and greater than 1. AKA is it a realistic value ##
+league_data$PowerPlay_Game > 1
+league_data$PowerPlay_Game < 6
+
+#### Skater Dataset Tests #### 
+
+## Let's test if forward points per game is less than 2 and greater than 1
+master_forwards$Points_Per_Game > 1
+master_forwards$Points_Per_Game < 2
+
+## Let's test if defenseman points per game is less than 1.5 and greater than 0.5
+master_dmen$Points_Per_Game > 0.5
+master_dmen$Points_Per_Game < 1.5

### 2024-04-15 scripts/04-model.R

@@ -1,0 +1,35 @@
+#### Preamble ####
+# Purpose: Generates a linear model based on NHL data
+# Author: Jason Ngo
+# Data: Apr 16 2023
+# Contact: jason_ngo@live.com
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+league_data <- read.csv(here::here("outputs/data/league_data.csv"))
+
+### Model data ####
+goals_model <-
+  stan_glm(
+    formula = Goals_PG ~ SV_Pct + GA_Avg + Sh_Pct + PP_PG * PP_Pct,
+    data = league_data,
+    family = gaussian(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 1122
+  )
+
+
+#### Save model ####
+saveRDS(
+  goals_model,
+  file = here::here("outputs/models/goals_model.rds")
+)
+
+

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/outputs/paper/paper.qmd
update

---
title: "Urban Connectivity Dynamics: Analyzing Free Public WiFi Distribution in Toronto"
subtitle: "By Building Typology, Population, and Socioeconomic Indicators"
author: 
  - Bernice Bao
thanks: "Code and data are available at: <https://open.toronto.ca/dataset/city-of-toronto-free-public-wifi/>, and <https://open.toronto.ca/dataset/ward-profiles-25-ward-model/>."
date: 23 January 2024
date-format: long
abstract: "In the era of increasing digital dependence, understanding the intricacies of urban internet connectivity is pivotal for fostering an effective and technologically equitable society. Public WiFi is a vital tool in modern society, offering internet access and supporting economic development, education, and social connections. In response to the increasing importance of internet connectivity in contemporary urban life, I investigate the distribution of free public WiFi in Toronto. This paper aims to provide key statistics on the distribution of public WiFi in Toronto and analyze the impact of building types, demographics, and economic influences on its distribution to gain as comprehensive a picture as possible of the role and impact of public WiFi in our lives."
format: pdf
fig-pos: "H"
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

#### Workspace set-up ####
library(tidyverse)
library(lubridate)
library(knitr)
library(palmerpenguins)
library(dplyr)
library(here)

#### Read in the data ####
cleaned_wifi_data <-
  read_csv(
    file = here("outputs/data/cleaned_wifi_data.csv"),
    show_col_types = FALSE
  )
#### Read in the data ####
cleaned_ward_data <-
  read_csv(
    file = here("outputs/data/cleaned_ward_data.csv"),
    show_col_types = FALSE
  )
```

# Introduction {#sec-intro}

In an era characterized by an ever-deepening reliance on digital technologies, the intricate tapestry of urban internet connectivity is emerging as a pivotal force in the transformation of our societal fabric. Countries globally are acknowledging internet access as an indispensable tool for active participation in contemporary democratic societies. In Canada, the Canadian Radio-television and Telecommunications Commission (CRTC) has recognized that "broadband internet access services are crucial to Canada’s economic, social, democratic, and cultural fabric."(@citiesfordigitalrights-toronto-wifi) As a university student navigating the complexities of contemporary living, the significance of understanding these dynamics cannot be overstated. We are acutely aware of the transformative potential that connectivity holds.

Public WiFi, often taken for granted in our daily lives, emerges as a vital tool shaping the contours of a technologically inclusive society. It plays a pivotal role in modern society, offering a myriad of benefits to individuals.  Beyond providing a cost-effective alternative to cellular data, it fosters internet accessibility, enabling people to stay connected without constraints.  This flexibility and mobility are particularly advantageous for students, travelers, and those on a budget, allowing them to work, study, or communicate from diverse locations.  Moreover, public WiFi contributes to social inclusion, bridging the digital divide and ensuring that even those without home internet access can engage in online activities.  Access to information, job searching, and remote work opportunities are further facilitated, enhancing the overall quality of life.  In emergencies, public WiFi serves as a crucial communication lifeline, demonstrating its indispensable role in connecting people and providing access to vital services.(@broadbandsearch-public-wifi)

Beyond the conveniences of internet access, it serves as a linchpin fostering economic development, facilitating education, and fortifying social connections. The present study, anchored in the evolving landscape of Toronto, endeavors to unravel the mysteries surrounding the distribution of public WiFi. This paper embarks on a compelling exploration of the distribution of free public WiFi in Toronto, recognizing it as not merely a technological amenity but a cornerstone of our modern existence. By examining the impact of various factors, such as building types, demographic patterns, and economic influences, this research seeks to paint a comprehensive portrait of the role and impact of public WiFi in our lives, providing readers with a lens to comprehend the symbiotic relationship between technology and urban living.

This research is divided into three main parts: `Introduction`(@sec-intro), `Data`(@sec-data), and `Conclusion`(@sec-conclusion). Commencing with the Data section, a meticulous exposition ensues, delineating the origin of datasets garnered from the OpenDataToronto Library and expounding upon the procedural rigors applied for data refinement and analysis. The conclusion part shows what I found during the analysis, while goes deeper into those findings, and finally wraps up the main discoveries from this paper about public free WiFi in Toronto.

# Data {#sec-data}

In this section, I will explore the diverse modes of `R`-based data collection and generation within the `RStudio` environment [@citeR; @citeRStudio]. The data under scrutiny in this paper originates from the `opendatatoronto` R package [@citeODT], specifically drawing from the Toronto fire incidents dataset. The analysis benefits from additional functionalities harnessed from various R packages such as `tidyverse` [@citeTidyverse], `here` [@citeHere], `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], and `dplyr` [@citeDplyr]. Further insights into the deployment of these packages will be expounded upon in the ensuing subsections.

## Location and description data of Toronto free public WiFi

The data I mainly used to generate the visualizations in this paper is City of Toronto Free Public WiFi from Open Data Toronto [@citeODT] that is an open source tool designed to be publicly available and encourage the development of valuable insights. The dataset, released by Toronto's Department of Information & Technology, captures information on location and description data free public WiFi locations in the City of Toronto and is updated monthly.

The data used for analysis in this article is as of January 3, 2024. According to the data characteristics described on the portal, this data should include information such as address, building type, building name, postal code, whether there is public wifi, etc. Through systematic analysis, it is found that only building type is the main subject variable that can be used to study the influencing factors of distribution. In order to quickly extract available data, only columns with reasonable data are retained during the data cleaning process (see Table 1). Basic data cleaning of column values to shorten the description of different types of clinics and improve readability. I create a statistic on the basis of monthly groups, using kable() from knitr to create @tbl-wifi.

```{r}
#| label: tbl-wifi
#| tbl-cap: Sample of cleaned Free Public WiFi data
#| echo: false

cleaned_wifi_data %>%
  head(10) %>%
  kable(
    col.names = c("Building type", "Ward", "Ward Number"),
    digits = 1
  )

```

From the Location and description data of Toronto free public WiFi, as of January 3, 2024, there are 194 free public WiFi location in the City of Toronto. Further breakdown shows that there are 27 free public WiFi location that are set in Arena, 74 free public WiFi location that are set in Community Recreation Centre, 17 free public WiFi location that are set in Outdoor Pool Building clinics, and  100 free public WiFi location that are set in library. Since the same wifi may cover multiple buildings, the sum of the number of public wifi corresponding to each building type is greater than the total number of public wifi in Toronto. The four building types mentioned above have the largest number of public wifi, covering almost 90% of total.

Library has the highest number of free public WiFi, which means that, in comparison to other types of buildings in Toronto, libraries offer the highest availability or coverage of free public WiFi services. This suggests that libraries prioritize or invest in providing accessible and free internet connectivity to the public within Toronto. Besides, Community Recreation Centre has the second most amount of free public WiFi. This information is indicative of the efforts or infrastructure investments made by Community Recreation Centres to offer accessible and free internet connectivity to the public.

## 2021 Ward Profiles based on the 25-Ward model

In order to deeply understand the factors affecting the distribution of public wifi in Toronto, I will study the population and income of each ward in Toronto on this basis. Therefore, the regional profile dataset (Data 2021), based on the 2021 Census data, was also included in the analysis. The dataset, published by City Planning, was last updated on January 3, 2024. The dataset contains demographic, social and economic information such as income and population for each ward.[@toronto-ward-profiles]

Ward profile data is stored in an Excel file with multiple tabs. The relevant data used for this analysis is included as a variable in the first TAB, the 2021 Census. Therefore, only the data from this TAB is downloaded for analysis. Further data cleansing was done to transform the data, retaining only information related to income and population levels in each ward (see @tbl-ward). Based on the Toronto Open Data Portal, the package has a CSV file, 25 ward names and numbers, which contains the mapping between ward codes and ward names. I create statistic to visualize these data, using kable() from knitr to create @tbl-ward.

```{r}
#| label: tbl-ward
#| tbl-cap: Sample of cleaned Ward of Toronto data
#| echo: false

cleaned_ward_data %>%
  head(10) %>%
  kable(
    col.names = c("Population", "Income", "Ward Number"),
    digits = 1
  )

```

There are 25 wards in the City of Toronto. Based 2021 census data, the wards with the highest population are:
Etobicoke-Lakeshore (Ward 3) at 139920, Spadina-Fort York (Ward 10) at 135400, and Etobicoke Centre (Ward 2) at 117200. The wards with the lowest population are: Scarborough North (Ward
23) at 94025, Don Valley East (Ward 16) at 94335, and Don Valley West (Ward
15) at 101025.

The wards with the highest household income are: Spadina-Fort York (Ward 10) at $ 80730,
Toronto Centre (Ward 13) at $68965, and Etobicoke-Lakeshore (Ward 3) at $65575. The
wards with the lowest household income are: Humber River-Black Creek (Ward 7) at $65,458,
York South-Weston (Ward 5) at $67,964, and Scarborough Centre (Ward 21) at $70,624.

# Conclusion {#sec-conclusion}

## Demand

**Societal need for public Wi-Fi differs according to building types**

The type of building can often reflect the different needs of modern people for WiFi, and having more public WiFi coverage proves that it has a stronger degree of Internet digitization needs.

For example, libraries have embraced technology, expanding their offerings beyond books to include a rich array of digital resources. Public WiFi is now a vital tool for students, researchers, and casual readers, facilitating access to online databases, ebooks, and collaborative group projects. Moreover, Community Recreation Centre have transformed into dynamic digital hubs, serving as venues for social gatherings and work. Here, individuals can check emails, work remotely, and stay informed about current events, all while enjoying a cup of coffee.

## Population

**Higher population density correlates with increased distribution of free public WiFi**

```{r}
#| label: fig-1
#| fig-cap: Relationship between Population of each Ward and Amount of Free Public WiFi Location
#| echo: false
#| warning: false
#| message: false

analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))

analysis_data_1 |> 
  ggplot(aes(x = Population, y = WardNumber)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Population of each Ward",
       y = "Ward Number")

```

From this graph, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.

In fact, this is similar to the distribution of free public WiFi in Toronto, but it does not exactly match. Ward 3, as the most populous area, has 13 public WiFi, which is a relatively large number in all Wards, and Ward 10 has 5, which is an insignificant number. But overall, for different Wards, a higher population slightly results in a higher amount of public WiFi.

## Income

**Income levels have limited influence on availability of free public Wi-Fi**

```{r}
#| label: fig-2
#| fig-cap: Relationship between Income of each Ward and Amount of Free Public WiFi Location
#| echo: false
#| warning: false
#| message: false

analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))

analysis_data_1 |> 
  ggplot(aes(x = Income, y = WardNumber)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Income of each Ward",
       y = "Ward Number")

```

From this graph, we can clearly see that Spadina-Fort York (Ward 10) has the greatest average income, Toronto Centre (Ward 13) has the second greatest average income, and Etobicoke-Lakeshore (Ward 3) has the third greatest average income.

We can conclude that there is no observable correlation between the number of Toronto public WiFi and the average household income in a ward.

\newpage

All in all, a significant number of people are increasingly relying on public Wi-Fi when they're out and about, underscoring their need for instant, hassle-free connectivity. Whether for work, social interaction or navigation, the modern lifestyle requires a seamless Internet experience, and the increasing number of academic and recreational projects joining the ranks of the application of public free WiFi is undoubtedly a strong support for the development of public WiFi in Toronto.

Furthermore, our data analysis indicates that while population size does exert a modest impact on the dissemination of free WiFi, the average income of a region lacks sufficient influence to be regarded as a determining factor in the allocation of public WiFi.

# References

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/00-simulate_data.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Get data on Location and description data of all City of Toronto free public WiFi locations and make table
+# Author: Bernice Bao
+# Date: 23 January 2024 
+# Contact: bernice.bao@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+
+
+#### Workspace setup ####
+library(tidyverse)
+install.packages("opendatatoronto")
+install.packages("knitr")
+
+library(knitr)
+library(janitor)
+library(lubridate)
+library(opendatatoronto)
+library(tidyverse)
+
+citation() # Get the citation information for R
+citation("ggplot2") # Get citation information for a package
+
+
+#### Simulate data ####
+set.seed(85)
+
+simulated_data <-
+  tibble(
+    has_wifi = rep(x = "Y", times = 100),
+    # Based on: https://www.toronto.ca/city-government/accountability-operations-customer-service/long-term-vision-plans-and-strategies/smart-cityto/internet-connectivity-connectto/connectto-free-public-wifi-locations/
+    building_type = c(
+      rep(x = "library", times = 20),
+      rep(x = "Arena", times = 40),
+      rep(x = "Community", times = 40)
+    ),
+    ward = rep(c(x = "Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"), times = 4),
+    ward_number =rep(c(1:25), times = 4)
+  )
+
+head(simulated_data)
+
+#### Save data ####
+write_csv(simulated_data,"outputs/data/simulated_data.csv")

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/01-download_data.R

@@ -1,0 +1,43 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the opendatatoronto r package
+# Author: Bernice Bao
+# Date: 23 January 2024
+# Contact: bernice.bao@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: NA
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+library(dplyr)
+
+#### Download data ####
+
+# get package
+package <- show_package("city-of-toronto-free-public-wifi")
+# get all resources for this package
+resources <- list_package_resources("city-of-toronto-free-public-wifi")
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# load the first datastore resource as a sample
+wifi_data <- filter(datastore_resources, row_number()==1) |> get_resource()
+wifi_data
+
+#### Save data ####
+# change raw_data to whatever name you assigned when you downloaded it.
+write_csv(wifi_data, "inputs/data/wifi_data.csv") 
+
+
+
+#### Download data 2####
+
+df = list_package_resources("6678e1a6-d25f-4dff-b2b7-aa8f042bc2eb") |>
+  filter(name == 
+           "2023-WardProfiles-2011-2021-CensusData") |>
+  get_resource()
+ward_data <- df$`2021 One Variable`
+
+#### Save data 2 ####
+write_csv(ward_data,"inputs/data/ward_data.csv")
+
+

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/02-data_cleaning.R

@@ -1,0 +1,71 @@
+#### Preamble ####
+# Purpose: Cleans the raw Toronto free public wifi data and 25 wards data from opendatatoronto r package
+# Author: Bernice Bao
+# Date: 23 January 2024
+# Contact: bernice.bao@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: run "01-download_data.R" before executing this file
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(dplyr)
+
+
+#### Clean data ####
+# load in raw data
+raw_wifi_data <- read_csv("inputs/data/wifi_data.csv")
+
+# clean and select column of interest
+cleaned_wifi_data <-
+  raw_wifi_data |>
+  filter(has_wifi == "Y")|> # filter on the locations with free public wifi
+  select(building_type, ward, ward_number) # only select the relevant columns
+  
+# rename the variables
+cleaned_wifi_data <-
+  cleaned_wifi_data |>
+  rename(BuildingType = building_type, Ward = ward, WardNumber = ward_number)
+
+head(cleaned_wifi_data)
+
+#### Save data ####
+# write cleaned data as csv
+write_csv(cleaned_wifi_data, "outputs/data/cleaned_wifi_data.csv")
+
+
+
+#### Clean data 2 ####
+# load in raw data
+raw_ward_data <- read_csv("inputs/data/ward_data.csv")
+
+# clean and select column of interest
+cleaned_ward_data <-
+  raw_ward_data |>
+  filter(`City of Toronto Profiles` == "Total - Age" | `City of Toronto Profiles` == "Total - Household total income groups in 2020 for private households - 25% sample data")|> # filter on the population and income
+  select(`City of Toronto Profiles`, ...3, ...4, ...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17, ...18, ...19, ...20, ...21, ...22, ...23, ...24, ...25, ...26, ...27) |># only select the relevant columns
+  t()
+
+# since this dataset organizes the data in first column as variables
+# the variables I interested in have to be considered as observations by R 
+# so I create a new data frame and keep all the data same 
+# move the variables to the correct position
+
+cleaned_ward_data <- data.frame(cleaned_ward_data)
+colnames(cleaned_ward_data ) <- cleaned_ward_data[1,]
[…]

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/03-test_data.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Tests if the simulated data is within the required bounds
+# Author: Bernice Bao
+# Date: 24 January 2023 
+# Contact: bernice.bao@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: run 00-simulate_data.R
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Test data ####
+
+data = read_csv("outputs/data/simulated_data.csv")
+
+# Test1
+all(data$ward_number >= 1 & data$ward_number <= 25)
+
+#Test2
+all(data$ward %in% c("Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"))

### 2024-04-16 outputs/paper/paper.qmd
revise the structure

@@
 ---
-title: "Better Talent and Rule Changes Have Resulted in More Goals in the NHL"
-subtitle: "An Analysis of the Increase in NHL Goal Scoring From 2011 to 2023"
+title: "Urban Connectivity Dynamics: Analyzing Free Public WiFi Distribution in Toronto"
+subtitle: "By Building Typology, Population, and Socioeconomic Indicators"
 author: 
-  - Jason Ngo
-thanks: "Code and data are available at: https://github.com/thejasonminh/nhlpoints"
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-format: 
-  pdf:
-    documentclass: article
-    geometry: margin = 1in
-    abstract: "Of the four major sports in the USA, hockey is the least popular.
-In an effort to grow popularity, the NHL (National Hockey League) strove to increase goal scoring.
-Using data from the NHL, this paper will analyze average leaguewide goal scoring per game from the 2011-2012 to 2022-2023 seaons.
-A linear regression model will also be created to predict goal scoring averages in the upcoming season.
-Results show that the NHL was succesful as the average goals per game increased over a 10 year period.
-These findings suggest that increased player talent and rule changes are primary factors."
+  - Bernice Bao
+thanks: "Code and data are available at: <https://github.com/BerniceBao/Free-Public-WiFi-Distribution-in-Toronto>
+ <https://open.toronto.ca/dataset/city-of-toronto-free-public-wifi/>, and <https://open.toronto.ca/dataset/ward-profiles-25-ward-model/>."
+date: 15 April 2024
+date-format: long
+abstract: "In the era of increasing digital dependence, understanding the intricacies of urban internet connectivity is pivotal for fostering an effective and technologically equitable society.
+Public WiFi is a vital tool in modern society, offering internet access and supporting economic development, education, and social connections.
+In response to the increasing importance of internet connectivity in contemporary urban life, I investigate the distribution of free public WiFi in Toronto.
+This paper aims to provide key statistics on the distribution of public WiFi in Toronto and analyze the impact of building types, demographics, and economic influences on its distribution to gain as comprehensive a picture as possible of the role and impact of public WiFi in our lives."
+format: pdf
+fig-pos: "H"
+toc: true
 number-sections: true
@@
 #| message: false
+#| echo: false
 
+#### Workspace set-up ####
 library(tidyverse)
+library(lubridate)
+library(knitr)
+library(palmerpenguins)
+library(dplyr)
 library(here)
-library(knitr)
 
-league_data <- read.csv(here::here("outputs/data/league_data.csv"))
-master_dmen <- read.csv(here::here("outputs/data/master_dmen.csv"))
-master_forwards <- read.csv(here::here("outputs/data/master_forwards.csv"))
-goals_model <- read_rds(here::here("outputs/models/goals_model.rds"))
+#### Read in the data ####
+cleaned_wifi_data <-
+  read_csv(
+    file = here("outputs/data/cleaned_wifi_data.csv"),
+    show_col_types = FALSE
+  )
+#### Read in the data ####
+cleaned_ward_data <-
+  read_csv(
+    file = here("outputs/data/cleaned_ward_data.csv"),
+    show_col_types = FALSE
+  )
 ```
 
-# Introduction
-Ice Hockey, Baseball, Football, and Basketball have become so popular in the United States, they've become branded as "The Big Four." The top leagues that operate in each sport: The NHL, MLB, NFL, and NBA are all centralized in the USA.
-Many of the top players that have played, or currently play, in these leagues are American, and their popularity only serves to inspire more youth to follow in their footsteps.
-There are no other sports that can threaten the stranglehold that these four sports have on the USA.
+# Introduction {#sec-intro}
 
-However, hockey has fallen behind the other three.
-Viewership, revenue and general popularity all pale in comparison to the other three sports [@citestatista].
-As a result; player salaries are lower in comparison to other sports and there is less youth talent in the pipelines.
-Two of the biggest factors for this decline are: High price of entry and low scoring.
-@citegear details the high upfront price of equipment and the amount of equipment required to play ice hockey.
-Because of this high upfront price, many kids elect not to choose hockey as their sport and instead elect for cheaper options such as basketball.
-Low scoring was a result of goalies having too much freedom, and teams creating bulletproof defensive tactics that involved clogging the middle of the ice and using their goalie to stifle momentum.
+In an era characterized by an ever-deepening reliance on digital technologies, the intricate tapestry of urban internet connectivity is emerging as a pivotal force in the transformation of our societal fabric.
+Countries globally are acknowledging internet access as an indispensable tool for active participation in contemporary democratic societies.
+In Canada, the Canadian Radio-television and Telecommunications Commission (CRTC) has recognized that "broadband internet access services are crucial to Canada’s economic, social, democratic, and cultural fabric."(@citiesfordigitalrights-toronto-wifi) As a university student navigating the complexities of contemporary living, the significance of understanding these dynamics cannot be overstated.
+We are acutely aware of the transformative potential that connectivity holds.
 
-As a response to the declining popularity, the NHL decided to implement rule changes and goalie equipment resizing to facilitate the increase of goal scoring.
-By increasing goal scoring, games would be more exciting to watch which in turn, would lead to increased viewership.
-This paper will analyze the effectiveness of these changes by drawing data from the NHL to analyze the estimand: Average leaguewide goals scored per game from the 2011-2012 season to the current (2022-2023) season.
-In addition, the paper will draw on a linear regression model to predict average goal scoring for the future season.
+Public WiFi, often taken for granted in our daily lives, emerges as a vital tool shaping the contours of a technologically inclusive society.
+It plays a pivotal role in modern society, offering a myriad of benefits to individuals.
+Beyond providing a cost-effective alternative to cellular data, it fosters internet accessibility, enabling people to stay connected without constraints.
+This flexibility and mobility are particularly advantageous for students, travelers, and those on a budget, allowing them to work, study, or communicate from diverse locations.
+Moreover, public WiFi contributes to social inclusion, bridging the digital divide and ensuring that even those without home internet access can engage in online activities.
+Access to information, job searching, and remote work opportunities are further facilitated, enhancing the overall quality of life.
+In emergencies, public WiFi serves as a crucial communication lifeline, demonstrating its indispensable role in connecting people and providing access to vital services.(@broadbandsearch-public-wifi)
 
-The findings show an increase of average goals scored per game over a 12 season period, with this current season exhibiting the highest goals scored per game during this period.
-The rule changes have given players more freedom to play and goalie equipment resizing has made goalies more susceptible to getting scored on.
-In addition, player talent has greatly increased over the years, with both forwards and defensemen exhibiting drastic offensive growth.
+Beyond the conveniences of internet access, it serves as a linchpin fostering economic development, facilitating education, and fortifying social connections.
+The present study, anchored in the evolving landscape of Toronto, endeavors to unravel the mysteries surrounding the distribution of public WiFi.
+This paper embarks on a compelling exploration of the distribution of free public WiFi in Toronto, recognizing it as not merely a technological amenity but a cornerstone of our modern existence.
+By examining the impact of various factors, such as building types, demographic patterns, and economic influences, this research seeks to paint a comprehensive portrait of the role and impact of public WiFi in our lives, providing readers with a lens to comprehend the symbiotic relationship between technology and urban living.
 
-In @sec-data of the paper, the source of the data as well as the three primary datasets are discussed.
-Strengths and weaknesses, methods of collection and terminologies will be explored.
-@sec-model of the paper will introduce a linear model, its justifications, and how it was constructed.
-In addition, the model will predict goal scoring averages for next season.
-@sec-results will discuss the results, touching on the trends and events that occurred over a 12-season period.
-@sec-discussion of the paper will touch on the factors that contributed to increased goal scoring as well as weaknesses and next steps.
+This research is divided into three main parts: `Introduction`(@sec-intro), `Data`(@sec-data), and `Conclusion`(@sec-conclusion).
+Commencing with the Data section, a meticulous exposition ensues, delineating the origin of datasets garnered from the OpenDataToronto Library and expounding upon the procedural rigors applied for data refinement and analysis.
+The conclusion part shows what I found during the analysis, while goes deeper into those findings, and finally wraps up the main discoveries from this paper about public free WiFi in Toronto.
 
 # Data {#sec-data}
-There are three primary datasets used for this paper.
-The first dataset, @sec-leagueavg, contains league averages from NHL seasons 2011-12 to 2022-23.
-The next dataset, @sec-fwds, contains the average points-per-game for every forward from the 2011-12 to 2022-23 seasons with a recorded average points-per-game pace of 1.00 or above.
-The last dataset, @sec-dmen, contains the average points-per-game for every defenseman from seasons 2011-12 to 2022-23 with a recorded average points-per-game pace of 0.60 or above.
 
-The analysis of these three datasets will be carried out using the statistical programming language `R` [@citeR], using the `tidyverse`[@citetidy], `here`[@citehere], and `readxl`[@citexl] packages.
-The figures and tables in the paper are generated using the, respectively, `ggplot2`[@citegg] and `knitr`[@citeknitr] packages.
+In this section, I will explore the diverse modes of `R`-based data collection and generation within the `RStudio` environment [@citeR; @citeRStudio].
+The data under scrutiny in this paper originates from the `opendatatoronto` R package [@citeODT], specifically drawing from the Toronto fire incidents dataset.
+The analysis benefits from additional functionalities harnessed from various R packages such as `tidyverse` [@citeTidyverse], `here` [@citeHere], `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], and `dplyr` [@citeDplyr].
+Further insights into the deployment of these packages will be expounded upon in the ensuing subsections.
 
-## League Averages Dataset {#sec-leagueavg}
+## Location and description data of Toronto free public WiFi
 
-The league averages dataset, represented below in @tbl-1, is created through averaging each variable across every team in the league for the corresponding season.
-Seasons 2011-12 to 2016-17 has averages for 30 teams, seasons 2017-18 to 2020-21 has averages for 31 teams, and seasons 2021-22 and 2022-23 has averages for 32 teams.
-Each team is responsible for tracking their own statistics, which is then conglomerated and represented by both the NHL and third party websites.
-Statistics tracking is done during real time, while a game is in progress.
-Internal models are used to calculate percentage, average and advanced statistics.
+The data I mainly used to generate the visualizations in this paper is City of Toronto Free Public WiFi from Open Data Toronto [@citeODT] that is an open source tool designed to be publicly available and encourage the development of valuable insights.
+The dataset, released by Toronto's Department of Information & Technology, captures information on location and description data free public WiFi locations in the City of Toronto and is updated monthly.
 
-With every team using their own methods for tracking statistics, measurement bias comes into play.
-Because the NHL does not standardize methods for tracking statistics, some teams will skew statistics such that their team is represented better.
+The data used for analysis in this article is as of January 3, 2024.
+According to the data characteristics described on the portal, this data should include information such as address, building type, building name, postal code, whether there is public wifi, etc.
+Through systematic analysis, it is found that only building type is the main subject variable that can be used to study the influencing factors of distribution.
+In order to quickly extract available data, only columns with reasonable data are retained during the data cleaning process (see Table 1).
+Basic data cleaning of column values to shorten the description of different types of clinics and improve readability.
+I create a statistic on the basis of monthly groups, using kable() from knitr to create @tbl-wifi.
 
 ```{r}
-#| label: tbl-1
-#| tbl-cap: The dataset containing averages from NHL seasons 2011-2012 to 2022-2023.
+#| label: tbl-wifi
+#| tbl-cap: Sample of cleaned Free Public WiFi data
+#| echo: false
+
+cleaned_wifi_data %>%
+  head(10) %>%
+  kable(
+    col.names = c("Building type", "Ward", "Ward Number"),
+    digits = 1
+  )
+
+```
+
+From the Location and description data of Toronto free public WiFi, as of January 3, 2024, there are 194 free public WiFi location in the City of Toronto.
+Further breakdown shows that there are 27 free public WiFi location that are set in Arena, 74 free public WiFi location that are set in Community Recreation Centre, 17 free public WiFi location that are set in Outdoor Pool Building clinics, and  100 free public WiFi location that are set in library.
+Since the same wifi may cover multiple buildings, the sum of the number of public wifi corresponding to each building type is greater than the total number of public wifi in Toronto.
+The four building types mentioned above have the largest number of public wifi, covering almost 90% of total.
+
+Library has the highest number of free public WiFi, which means that, in comparison to other types of buildings in Toronto, libraries offer the highest availability or coverage of free public WiFi services.
+This suggests that libraries prioritize or invest in providing accessible and free internet connectivity to the public within Toronto.
+Besides, Community Recreation Centre has the second most amount of free public WiFi.
+This information is indicative of the efforts or infrastructure investments made by Community Recreation Centres to offer accessible and free internet connectivity to the public.
+
+## 2021 Ward Profiles based on the 25-Ward model
+
+In order to deeply understand the factors affecting the distribution of public wifi in Toronto, I will study the population and income of each ward in Toronto on this basis.
+Therefore, the regional profile dataset (Data 2021), based on the 2021 Census data, was also included in the analysis.
+The dataset, published by City Planning, was last updated on January 3, 2024.
+The dataset contains demographic, social and economic information such as income and population for each ward.[@toronto-ward-profiles]
+
+Ward profile data is stored in an Excel file with multiple tabs.
+The relevant data used for this analysis is included as a variable in the first TAB, the 2021 Census.
+Therefore, only the data from this TAB is downloaded for analysis.
+Further data cleansing was done to transform the data, retaining only information related to income and population levels in each ward (see @tbl-ward).
+Based on the Toronto Open Data Portal, the package has a CSV file, 25 ward names and numbers, which contains the mapping between ward codes and ward names.
+I create statistic to visualize these data, using kable() from knitr to create @tbl-ward.
+
+```{r}
+#| label: tbl-ward
+#| tbl-cap: Sample of cleaned Ward of Toronto data
+#| echo: false
+
+cleaned_ward_data %>%
+  head(10) %>%
+  kable(
+    col.names = c("Population", "Income", "Ward Number"),
+    digits = 1
+  )
+
+```
+
+There are 25 wards in the City of Toronto. Based 2021 census data, the wards with the highest population are:
+Etobicoke-Lakeshore (Ward 3) at 139920, Spadina-Fort York (Ward 10) at 135400, and Etobicoke Centre (Ward 2) at 117200. The wards with the lowest population are: Scarborough North (Ward
+23) at 94025, Don Valley East (Ward 16) at 94335, and Don Valley West (Ward
+15) at 101025.
+
+The wards with the highest household income are: Spadina-Fort York (Ward 10) at $ 80730,
+Toronto Centre (Ward 13) at $68965, and Etobicoke-Lakeshore (Ward 3) at $65575. The
+wards with the lowest household income are: Humber River-Black Creek (Ward 7) at $65,458,
+York South-Weston (Ward 5) at $67,964, and Scarborough Centre (Ward 21) at $70,624.
+
+# Conclusion {#sec-conclusion}
+
+## Demand
+
+**Societal need for public Wi-Fi differs according to building types**
+
+The type of building can often reflect the different needs of modern people for WiFi, and having more public WiFi coverage proves that it has a stronger degree of Internet digitization needs.
+
+For example, libraries have embraced technology, expanding their offerings beyond books to include a rich array of digital resources.
+Public WiFi is now a vital tool for students, researchers, and casual readers, facilitating access to online databases, ebooks, and collaborative group projects.
+Moreover, Community Recreation Centre have transformed into dynamic digital hubs, serving as venues for social gatherings and work.
+Here, individuals can check emails, work remotely, and stay informed about current events, all while enjoying a cup of coffee.
+
+## Population
+
+**Higher population density correlates with increased distribution of free public WiFi**
+
+```{r}
+#| label: fig-1
+#| fig-cap: Relationship between Population of each Ward and Amount of Free Public WiFi Location
 #| echo: false
 #| warning: false
+#| message: false
 
-kable(league_data, padding = 5)
+analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))
+
+analysis_data_1 |> 
+  ggplot(aes(x = Population, y = WardNumber)) +
+  geom_point(alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Population of each Ward",
+       y = "Ward Number")
+
 ```
 
-Almost all of @tbl-1 is taken from @citehref, with the exception of the `Sh_Pct` column, which was pulled from @citenhl and merged with the larger dataset.
-The reasoning why I didn't use the NHL website for the majority of this dataset was that the NHL website did not support automatic calculation of league averages.
-There are 8 columns total in this league averages dataset:
+From this graph, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.
 
-1. The `Season` column, which denotes the NHL season each average represents. 
-2. The `Goals_PG` column, which denotes the average goals per game for the respective season. 
-3. The `PP_PG` column, which denotes the average power plays, situations where one team has a man-advantage due to a penalty, per game for the respective season. 
-4. The `PP_Pct` column, which shows the percentage of power plays that were successful, or resulted in a goal, for the respective season. 
-5. The `SA` column, which details the leaguewide average shots a team faced per game, for the respective season.
-6. The `SV_Pct` column, which denotes the leaguewide average save percentage, which represents the percentage of total saves made, per season.
-7. The `GA_Avg` column, which represents the leaguewide average goals against per season.
-8. The `Sh_Pct` columnm which represents the leaguewide average shooting percentage. Shooting percentage represents the percentage of shots that are goals.
+In fact, this is similar to the distribution of free public WiFi in Toronto, but it does not exactly match.
+Ward 3, as the most populous area, has 13 public WiFi, which is a relatively large number in all Wards, and Ward 10 has 5, which is an insignificant number.
+But overall, for different Wards, a higher population slightly results in a higher amount of public WiFi.
 
-The `Season` column was necessary as a timeframe is needed to anchor this analysis.
-The `Goals_PG` was chosen as it is the estimand of this paper.
-The remaining 6 columns were chosen as they all correlate to goal scoring, or situations that incentivizes goal scoring.
-For example, having a high shooting percentage correlates to having a higher goals per game, which would decrease the save percentage and increase the goals against average.
-The powerplay columns represent the opportunities teams have to score on the man advantage and the capability for teams to score on that advantage.
+## Income
 
-## Forwards Dataset {#sec-fwds}
-
-The forwards dataset, represented below in @tbl-2, represents the count of every forward from the 2011-12 to 2022-23 seasons that finished the season with, at minimum, a 1 point-per-game (PPG) pace.
-A point represents either a goal or an assist.
-The reason why I chose to filter the forwards by the 1 PPG pace is because there are few players that finish the season at a 1 PPG pace, which makes those players that are able to do so offensively gifted.
-
-The calculation of average points-per-game is quite simple: Simply take a player's total points and ratio it against the amount of games they've played.
-It is very simple to track a player's points since every game log has every goal with their respective assists recorded.
-This is true for both forwards and defensemen.
+**Income levels have limited influence on availability of free public Wi-Fi**
 
 ```{r}
-#| label: tbl-2
-#| tbl-cap: The count of forwards that finish at a 1 PPG pace or above per season.
+#| label: fig-2
+#| fig-cap: Relationship between Income of each Ward and Amount of Free Public WiFi Location
 #| echo: false
 #| warning: false
+#| message: false
 
-kable(master_forwards |> count(Season) |> rename("PPG Players" = "n"), padding = 5)
+analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))
+
+analysis_data_1 |> 
+  ggplot(aes(x = Income, y = WardNumber)) +
+  geom_point(alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Income of each Ward",
+       y = "Ward Number")
+
 ```
 
-All the data about forwards was taken from [@citenhl]. Since we are only concerned with the average offensive output of forwards, I chose to only analyze the points-per-game pace of forwards.
+From this graph, we can clearly see that Spadina-Fort York (Ward 10) has the greatest average income, Toronto Centre (Ward 13) has the second greatest average income, and Etobicoke-Lakeshore (Ward 3) has the third greatest average income.
 
-## Defensemen Dataset {#sec-dmen}
-
-The defensemen dataset, represented below in @tbl-3, represents the count of every defensemen from the 2011-12 to 2022-23 seasons that finished the season with, at minimum, a 0.6 point-per-game pace.
-Rather than filtering by the 1 PPG pace like I did for forwards, I chose to filter defensemen by a 0.6 PPG pace.
-The reasoning for doing so is because defensemen simply do not contribute offensively as much as forwards do.
-A defensemen will rarely match a forward's offensive output because offense isn't the primary objective of a defensemen.
-However, talented defensemen that are table to contribute offensively usually do so at a 0.6 point-per-game pace or above.
-
-```{r}
-#| label: tbl-3
-#| tbl-cap: The count of defensemen that finish at a 0.6 PPG pace or above per season.
-#| echo: false
-#| warning: false
-
-kable(master_dmen |> count(Season) |> rename("PPG Players" = "n"), padding = 5)
-```
-
-All the data regarding defensemen was also taken from [@citenhl].
-Similar to the forwards, I am only concerned with the offensive output of defensemen so I chose to only analyze the point-per-game pace of defensemen.
-
-# Model {#sec-model}
-
-The goal of my modelling strategy is twofold.
-Firstly, I want to analyze to what effect power play opportunities, power play effectiveness, save percentage, goals against average and shooting percentages all have on average goals scored per game.
-Secondly, I want to use these results to predict the average goals scored per game for the next season.
-
-In the section below, the Bayesian analysis model used to investigate the multiple linear regression model of average goals scored per game is detailed.
-
-## Model set-up
-
-Define $y_i$ as the leaguewide average goals scored per game.
-Then $\beta_1$ is the leaguewide average power plays per game, $\gamma_1$ is leaguewide powerplay percentage, $\theta_1$ is leaguewide save percentage, $\phi_1$ is the leaguewide goals against average, and $\kappa_1$ is the leaguewide average shooting percentage.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \theta_i + \phi_i + \kappa_i + \beta_i * \gamma_i \\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\theta &\sim \mbox{Normal}(0, 2.5) \\
-\phi &\sim \mbox{Normal}(0, 2.5) \\
-\kappa &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in `R` [@citeR] using the `rstanarm` package of [@citestan]. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between average goals scored per game and power play opportunities, power play percentage, goals against average, and shooting percentage.
-The higher those four variables are, the more goals there should be.
-In particular, there should also be a negative relationship between goals scored and save percentage.
-With a lower save percentage, there are less saves that are occuring which means there should be more goals.
-Additionally, the two power play variables are interacted with eachother, since they are directly related to one another.
-There cannot be a power play percentage without a power play opportunity.
-
-## Model Prediction
-Taking the differences in averages from the 2021-2022 season and the 2022-2023 season, I will create a test dataset for the 2023-2024 season by adding those differences onto the averages from the 2022-2023 season and use the model with this test dataset to predict the average goals scored per game for the 2023-2024 season.
-@tbl-4 is the resulting mock 2023-2024 season averages data.
-```{r}
-#| echo: false
-#| warning: false
-
-test_league <- tibble(
-  Season = c(
-    "2023-24"
-    ),
-  Goals_PG = c(
-    3.22
-  ),
-  PP_PG = c(
-    3.25
-  ),
-  PP_Pct = c(
-    22.01
-  ),
-  SA = c(
-    30.8
-  ),
-  SV_Pct = c(
-    0.901
-  ),
-  GA_Avg = c(
-    3.02
-  ),
-  Sh_Pct = c(
-    10.33
-  )
-)
-
-prediction <- predict(goals_model, newdata = test_league)
-```
-
-```{r}
-#| label: tbl-4
-#| tbl-cap: A mock league averages dataset for the 2023-2024 NHL season.
-#| echo: false
-#| warning: false
-
-kable(test_league, padding = 5)
-```
-
-Using the mock data alongside the model, the predicted average goals per game for the 2023-2024 season is **3.21**.
+We can conclude that there is no observable correlation between the number of Toronto public WiFi and the average household income in a ward.
 
@@
 
-# Results {#sec-results}
+All in all, a significant number of people are increasingly relying on public Wi-Fi when they're out and about, underscoring their need for instant, hassle-free connectivity.
+Whether for work, social interaction or navigation, the modern lifestyle requires a seamless Internet experience, and the increasing number of academic and recreational projects joining the ranks of the application of public free WiFi is undoubtedly a strong support for the development of public WiFi in Toronto.
 
-## Goals, Saves and Shooting
-
-```{r}
-#| label: fig-1
-#| fig-cap: An increase of average goals scored per game over 12 seasons.
-#| echo: false
-#| warning: false
-
-library(tidyverse)
-
-league_data |> ggplot(aes(x = Season, y = Goals_PG)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Leaguewide Average Goals Scored Per Game over 12 Seasons",
-    x = "Season",
-    y = "Average Goals P/G"
-  ) 
-```
-
-@fig-1 shows an upward trajectory of goal scoring per season.
-The horizontal axis denotes the season, and the vertical axis denotes the average goals scored per game.
-There are two things of note: First, there was a big spike in goal scoring starting in the 2017-18 season, with an increase of 0.2 from the 2016-17 season.
-with the subsequent seasons building on that jump.
-The second thing of note however, is the slight dip in scoring during the 2020-21 season where the average fell to 2.94 from the 3.02 the season before.
-
-In regards to the spike of scoring in the 2017-18 season, the NHL had implemented a few rule changes that helped facilitate scoring.
-Additionally, goalie equipment had shrunken by the time the season started, so goalies were scored on more.
-
-With respect to the slight dip in scoring in the 2020-21 season, that season took place during the height of the COVID-19 pandemic and due to health regulations, many players had to miss extended periods of time to self-quarantine after a positive test.
-The loss of skilled players for periods of time and the erosion of team chemistry led to the slight decrease in goals scored that season.
-
-\newpage
-
-```{r}
-#| label: fig-2
-#| fig-cap: Implications on increased goal scoring on save percentage, goals against average and shooting percentage.
-#| fig-subcap: ["Save Percentage", "Goals Against Average", "Shooting Percentage"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-library(tidyverse)
-
-league_data |> ggplot(aes(x = Season, y = SV_Pct)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Leaguewide Save Percentage over 12 seasons",
-    x = "Season",
-    y = "Average Save %"
-  )
-
-league_data |> ggplot(aes(x = Season, y = GA_Avg)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Leaguewide Goals Against Average over 12 seasons",
-    x = "Season",
-    y = "GAA"
-  )
-
-league_data |> ggplot(aes(x = Season, y = Sh_Pct)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Leaguewide Average Shooting Percentage over 12 seasons",
-    x = "Season",
-    y = "Average Sh %"
-  ) 
-```
-
-@fig-2 shows the implications of increased goal scoring in average save percentage, goals against average and average shooting percentage.
-A better shooting percentage means more goals are scored.
-More goals scored increases the goals against average, and more goals scored means less saves are made.
-
-\newpage
-
-## More Talent, Better Power Plays
-
-```{r}
-#| label: fig-3
-#| fig-cap: The rise of offensively gifted forwards and defensemen over 12 seasons.
-#| fig-subcap: ["Forwards", "Defensemen", "Shots Per Game"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-library(tidyverse)
-
-master_forwards |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Productive Forwards over 12 seasons",
-    x = "Season",
-    y = ">= 1 PPG Forwards"
-  )
-
-master_dmen |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Productive Defensemen over 12 seasons",
-    x = "Season",
-    y = ">= 0.6 PPG Defensemen"
-  )
-
-league_data |> ggplot(aes(x = Season, y = SA)) +
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Avg. Shots Per Game over 12 seasons",
-    x = "Season",
-    y = "Shots Per Game"
-  ) 
-```
-
-@fig-3 showcases the increase of offensively talented forwards and defensemen over 12 seasons, with an additional graph of leaguewide shots per game to contextualize the increase in talent.
-For offensively talented forwards, the horizontal axis denotes the season while the vertical axis denotes the amount of forwards that finished the season at a pace of 1 PPG or greater.
-Similarly, the offensively talented defensemen graph has the same horizontal axis, while the vertical axis denotes the amount of defensemen that finished the season at a pace of 0.6 PPG or greater.
-For shots per game, the horizontal axis once again denotes the season and the vertical axis denotes the average leaguewide shots per game.
-
-Similar to the above section about goal scoring, the shift from the 2016-17 to the 2017-18 season brought an uptick in scoring alongside it.
-Regarding forwards, it is interesting to note that 2012-13 total of 20 players was never reached again until the 2017-18 season.
-The 2014-15, 2015-16, and 2016-17 seasons all saw a drop in productive forwards.
-This trend is not reflected in the defensemen graph.
-Aside from a slight drop in the 2016-17 season, the amount of offensively productive defensemen slowly increased over the years.
-This trend means that over the seasons, defensemen have gotten more involved with the offense.
-This could be a result of increased talent, or a shifting in tactics to allow defensemen to flourish offensively, or perhaps a mixture of both.
-
-However, what makes this increase in productivity even more substantial is looking at the average shots per game over 12 seasons.
-Over the 12 season period, the lowest average shots per game was 29, and thee highest was 31.8.
-This means that teams are barely shooting more; the average shots per game over 12 seasons has remained somewhat consistent.
-Players are simply doing more with their shots on goal.
-
-\newpage
-
-```{r}
-#| label: fig-4
-#| fig-cap: The implications of better talent on power play effectiveness.
-#| fig-subcap: ["Amount of Power Plays", "Effectiveness of Power Plays"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-library(tidyverse)
-
-league_data |> ggplot(aes(x = Season, y = PP_PG)) + 
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Power Play Opportunities over 12 seasons",
-    x = "Season",
-    y = "Avg. PP Opportunities"
-  )
-
-league_data |> ggplot(aes(x = Season, y = PP_Pct)) + 
-  geom_line(group = 1) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Power Play Effectiveness over 12 seasons",
-    x = "Season",
-    y = "Avg. PP %"
-  ) 
-```
-
-One consequence of increased offensive talent is a greater effectiveness of the power play.
-@fig-4 shows that, despite a decrease in overall power plays over a 12 season period, power play effectiveness has trended upwards.
-In essence, players are making more with less, and are doing so because of how much more skilled players are in comparison to players before.
-
-# Discussion {#sec-discussion}
-
-As evidenced by the results, goal scoring has steadily increased over the seasons.
-The NHL was successful in their goal of making the game more exciting, but how did they go about doing it?
-The two primary factors that led to this increase was an increase of mid-to-high tier player talent, as well as rule changes.
-
-\newpage
-
-## Good Players Becoming Great
-
-While I talked about the increase of the amount of productive forwards and defensemen, here I will discuss the increase in the amount of good forwards and compare them to the amount of elite forwards.
-To be classified as an elite forward, usually they would need to finish the season at around a 1.4 PPG pace.
-Because of how hard of a feat it is to achieve a sustained 1.4 PPG pace, there are very few elite forwards in the NHL.
-@fig-5 shows the amount of forwards that finish in multiple PPG ranges.
-
-```{r}
-#| label: fig-5
-#| fig-cap: Different Forward PPG brackets over 12 seasons.
-#| fig-subcap: ["1.4 or above", "Between 1.2 and 1.4", "Between 1.0 and 1.2"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-library(tidyverse)
-
-master_forwards |> filter(Points_PG >= 1.4) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Forwards above 1.4 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  )
-
-master_forwards |> filter(Points_PG >= 1.2 & Points_PG < 1.4) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Forwards Between 1.2 and 1.4 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  )
-
-master_forwards |> filter(Points_PG >= 1.0 & Points_PG < 1.2) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Forwards Between 1.0 and 1.2 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  ) 
-```
-
-As evidenced, the amount of elite players have remained consistent over the seasons, although the big spike in the 2021-22 season may be a sign that times are changing.
-The bigger story here though, is the amount of players that finish in the 1.2 and 1.4 range from 2017-18 beyond.
-There is a dramatic spike in these productive players from before 2017-18 to that season and beyond.
-The graphs here show that there while the ceiling of player talent has remained the same, the floor is steadily increasing.
-
-\newpage
-
-The same level of growth can be seen with defensemen as well. @fig-6 shows the amount of defensemen that finish in multiple PPG brackets.
-
-```{r}
-#| label: fig-6
-#| fig-cap: Different Defenseman PPG brackets over 12 seasons.
-#| fig-subcap: ["0.95 or above", "Between 0.75 and 0.95", "Between 0.6 and 0.75"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-library(tidyverse)
-
-master_dmen |> filter(Points_PG >= 0.95) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Defensemen above 0.95 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  )
-
-master_dmen |> filter(Points_PG >= 0.75 & Points_PG < 0.95) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Defensemen Between 0.75 and 0.95 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  )
-
-master_dmen |> filter(Points_PG >= 0.6 & Points_PG < 0.75) |> count(Season) |> ggplot(aes(x = Season, y = n)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
-  labs(
-    title = "Defensemen Between 0.6 and 0.75 PPG over 12 seasons",
-    x = "Season",
-    y = "Player Count"
-  ) 
-```
-There is a slightly different trajectory with defensemen.
-It appears that while the floor of offensive production has remained somewhat consistent over the seasons, the ceiling has only gotten higher.
-It appears that we may be in the middle of a renaissance of offensively gifted defensemen, with multiple defensemen over the last two seasons exhibiting a PPG of 0.95 or above.
-To contextualize, a great forward should finish at a 1.00 PPG, so to see multiple defensemen reach those numbers is unprecedented.
-
-The floor of talent for forwards have increased, and the ceiling of talent for defensemen have increased, and both of these factors contributed to a spike in goal scoring.
-
-\newpage
-
-## Changes in 2017-18
-
-While the findings have shown a general increase in goal scoring, the biggest spike in productivity lies between the 2016-17 season and the 2017-18 season.
-While blossoming player talent is certainly a factor here, the biggest factor in play are rule changes and goalie equipment changes.
-
-### Rule Changes
-
-@citerules details all of the rule changes effective for the beginning of the 2017-18 season.
-These changes include a power play consequence for a failing an off-side challenge (forcing an analyzed replay to see if a play was offside), denying time-outs for teams committing icing (throwing the puck down the entire length of the ice to delay the game and to change shifts), face-off placement changes and stricter enforcement on the stick-slashing around a player's hands.
-All of these changes, save for the face-off placement changes, had an effect on goal scoring.
-The off-side challenge rule change introduced more power plays, denying time-outs prevents tired players from resting up during icing infractions which led to worse defense, and the enforcement of slashing led to more offensive freedom and creativity for players skilled at working with the puck.
-
-### Goalie Equipment Changes
-
-@citegoalies shows the timeline of NHL goalie equipment changes.
-These changes included making goalie equipment and padding smaller to facilitate more scoring.
-While the changes were put into effect during the latter half of the 2016-17 season, teams became accustomed to the adjusted size and found more ways to score on these smaller goalies in the 2017-18 season.
-This is shown in the increase of goals against average and the drop in save percentage.
-
-## Weaknesses
-
-The primary weakness of my analysis is how surface level it is.
-I only focused on basic statistics such as per-game and per-season averages.
-A more richer and defined analysis would've been done if advanced statistics were used in this analysis.
-Analysis on these advanced stats would've been conducted on a team level and on an individual level to further investigate the causes for the increase in goal scoring.
-
-Another weakness of my analysis is that it covers leaguewide averages per season.
-One problem with taking leaguewide averages is that in the NHL, there is very wide gap in skill between the teams at the top and teams at the bottom.
-Due to how drafting works in the NHL, teams at the bottom of the standings are encouraged to play worse in hopes of a better chance at drafting better players.
-In an analysis of leaguewide goal scoring, it feels wrong to include teams that are prioritizing losing over winning for a better future.
-
-## Next Steps
-
-An appropriate next step of this analysis is to compare the increase in goal scoring with viewership numbers, to see if the NHL was successful in drawing more eyes to the game.
-Additionally, to further analyze viewership rates, analysis on streaming options, media narratives and rivalry match-ups is recommended.
-Alongside the analysis on viewership, analysis on the amount of youth talent based on national viewership in not just America, but for European countries as well, is recommended to determine the outlook of the future of the NHL.
-
-\newpage
+Furthermore, our data analysis indicates that while population size does exert a modest impact on the dissemination of free WiFi, the average income of a region lacks sufficient influence to be regarded as a determining factor in the allocation of public WiFi.
 

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/00-simulate_data.R

@@ -1,44 +1,0 @@
-#### Preamble ####
-# Purpose: Get data on Location and description data of all City of Toronto free public WiFi locations and make table
-# Author: Bernice Bao
-# Date: 23 January 2024 
-# Contact: bernice.bao@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: NA
-
-
-#### Workspace setup ####
-library(tidyverse)
-install.packages("opendatatoronto")
-install.packages("knitr")
-
-library(knitr)
-library(janitor)
-library(lubridate)
-library(opendatatoronto)
-library(tidyverse)
-
-citation() # Get the citation information for R
-citation("ggplot2") # Get citation information for a package
-
-
-#### Simulate data ####
-set.seed(85)
-
-simulated_data <-
-  tibble(
-    has_wifi = rep(x = "Y", times = 100),
-    # Based on: https://www.toronto.ca/city-government/accountability-operations-customer-service/long-term-vision-plans-and-strategies/smart-cityto/internet-connectivity-connectto/connectto-free-public-wifi-locations/
-    building_type = c(
-      rep(x = "library", times = 20),
-      rep(x = "Arena", times = 40),
-      rep(x = "Community", times = 40)
-    ),
-    ward = rep(c(x = "Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"), times = 4),
-    ward_number =rep(c(1:25), times = 4)
-  )
-
-head(simulated_data)
-
-#### Save data ####
-write_csv(simulated_data,"outputs/data/simulated_data.csv")

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/01-download_data.R

@@ -1,43 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from the opendatatoronto r package
-# Author: Bernice Bao
-# Date: 23 January 2024
-# Contact: bernice.bao@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: NA
-
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-library(dplyr)
-
-#### Download data ####
-
-# get package
-package <- show_package("city-of-toronto-free-public-wifi")
-# get all resources for this package
-resources <- list_package_resources("city-of-toronto-free-public-wifi")
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-# load the first datastore resource as a sample
-wifi_data <- filter(datastore_resources, row_number()==1) |> get_resource()
-wifi_data
-
-#### Save data ####
-# change raw_data to whatever name you assigned when you downloaded it.
-write_csv(wifi_data, "inputs/data/wifi_data.csv") 
-
-
-
-#### Download data 2####
-
-df = list_package_resources("6678e1a6-d25f-4dff-b2b7-aa8f042bc2eb") |>
-  filter(name == 
-           "2023-WardProfiles-2011-2021-CensusData") |>
-  get_resource()
-ward_data <- df$`2021 One Variable`
-
-#### Save data 2 ####
-write_csv(ward_data,"inputs/data/ward_data.csv")
-
-

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/02-data_cleaning.R

@@ -1,71 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw Toronto free public wifi data and 25 wards data from opendatatoronto r package
-# Author: Bernice Bao
-# Date: 23 January 2024
-# Contact: bernice.bao@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: run "01-download_data.R" before executing this file
-
-#### Workspace setup ####
-library(tidyverse)
-library(janitor)
-library(dplyr)
-
-
-#### Clean data ####
-# load in raw data
-raw_wifi_data <- read_csv("inputs/data/wifi_data.csv")
-
-# clean and select column of interest
-cleaned_wifi_data <-
-  raw_wifi_data |>
-  filter(has_wifi == "Y")|> # filter on the locations with free public wifi
-  select(building_type, ward, ward_number) # only select the relevant columns
-  
-# rename the variables
-cleaned_wifi_data <-
-  cleaned_wifi_data |>
-  rename(BuildingType = building_type, Ward = ward, WardNumber = ward_number)
-
-head(cleaned_wifi_data)
-
-#### Save data ####
-# write cleaned data as csv
-write_csv(cleaned_wifi_data, "outputs/data/cleaned_wifi_data.csv")
-
-
-
-#### Clean data 2 ####
-# load in raw data
-raw_ward_data <- read_csv("inputs/data/ward_data.csv")
-
-# clean and select column of interest
-cleaned_ward_data <-
-  raw_ward_data |>
-  filter(`City of Toronto Profiles` == "Total - Age" | `City of Toronto Profiles` == "Total - Household total income groups in 2020 for private households - 25% sample data")|> # filter on the population and income
-  select(`City of Toronto Profiles`, ...3, ...4, ...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17, ...18, ...19, ...20, ...21, ...22, ...23, ...24, ...25, ...26, ...27) |># only select the relevant columns
-  t()
-
-# since this dataset organizes the data in first column as variables
-# the variables I interested in have to be considered as observations by R 
-# so I create a new data frame and keep all the data same 
-# move the variables to the correct position
-
-cleaned_ward_data <- data.frame(cleaned_ward_data)
-colnames(cleaned_ward_data ) <- cleaned_ward_data[1,]
[…]

### 2024-04-16 City-of-Toronto-Free-Public-WiFi-main/scripts/03-test_data.R

@@ -1,20 +1,0 @@
-#### Preamble ####
-# Purpose: Tests if the simulated data is within the required bounds
-# Author: Bernice Bao
-# Date: 24 January 2023 
-# Contact: bernice.bao@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: run 00-simulate_data.R
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Test data ####
-
-data = read_csv("outputs/data/simulated_data.csv")
-
-# Test1
-all(data$ward_number >= 1 & data$ward_number <= 25)
-
-#Test2
-all(data$ward %in% c("Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"))

### 2024-04-16 scripts/00-simulate_data.R

@@ -1,57 +1,44 @@
 #### Preamble ####
-# Purpose: Simulates the primary datasets of interest
-# Author: Jason Ngo
-# Data: Apr 16 2023
-# Contact: jason_ngo@live.com
+# Purpose: Get data on Location and description data of all City of Toronto free public WiFi locations and make table
+# Author: Bernice Bao
+# Date: 23 January 2024 
+# Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Some knowledge of NHL statistics terminology
+# Pre-requisites: NA
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(2223)
+install.packages("opendatatoronto")
+install.packages("knitr")
+
+library(knitr)
+library(janitor)
+library(lubridate)
+library(opendatatoronto)
+library(tidyverse)
+
+citation() # Get the citation information for R
+citation("ggplot2") # Get citation information for a package
+
 
 #### Simulate data ####
-## Season column correlates to the NHL season. Column XX represents season 20XX - 20YY where YY is (XX  + 1) ##
-## For example, column 16 represents season 2016 - 2017
+set.seed(85)
 
-### Tibble for the league data simulation ###
-## Avg_Goals_Game has a min of 1, since realistically every team should average 1 goal a game atleast ##
-## Avg_Shots_Game has a min of 20 and max of 45, since those averages are hard to come by in a normal NHL season
-## Avg_PP_Percentage has a min of 5 and a max of 40, for the same reasons above
-## Avg_Sh_Percentage has a min of 5 and a max of 20, for the same reasons above
-nhlleaguesim <- tibble(
-  Season = c(
-    11:22
-  ),
-  Avg_Goals_Game = c(
-    runif(n = 12, min = 1, max = 5)
-  ),
-  Avg_Shots_Game = c(
-    runif(n = 12, min = 20, max = 45)
-  ),
-  Avg_PP_Percentage = c(
-    runif(n = 12, min = 5, max = 40)
-  ),
-  Avg_Sh_Percentage = c(
-    runif(n = 12, min = 5, max = 20)
-  ),
-  Avg_PP_Opportunities = c(
-    runif(n = 12, min = 1, max = 6)
+simulated_data <-
+  tibble(
+    has_wifi = rep(x = "Y", times = 100),
[…]

### 2024-04-16 scripts/01-download_data.R

@@ -1,22 +1,43 @@
 #### Preamble ####
-# Purpose: Describes the acquisition process of the datasets used
-# Author: Jason Ngo
-# Data: Apr 16 2023
-# Contact: jason_ngo@live.com
+# Purpose: Downloads and saves the data from the opendatatoronto r package
+# Author: Bernice Bao
+# Date: 23 January 2024
+# Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: A working internet connection
-
+# Pre-requisites: NA
 
 #### Workspace setup ####
+library(opendatatoronto)
 library(tidyverse)
+library(dplyr)
 
 #### Download data ####
 
-## League averages data downloaded from https://www.hockey-reference.com/leagues/stats.html#stats ##
-## Data on shooting is from nhl.com/stats/teams, from seasons 2011-2012 to 2022-2023 and viewing Shots By Type report ##
-## Data on forwards and dmen is from nhl.com/stats/skaters, from seasons 2011-2012 to 2022-2023 with 15 minimum games ## ##
-## forwards data was selected by filtering all players with 1.0 points per game or above from those seasons ##
-## defensemen data was selected by filtering all players with 0.6 points per game or above from those seasons ##
+# get package
+package <- show_package("city-of-toronto-free-public-wifi")
+# get all resources for this package
+resources <- list_package_resources("city-of-toronto-free-public-wifi")
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# load the first datastore resource as a sample
+wifi_data <- filter(datastore_resources, row_number()==1) |> get_resource()
+wifi_data
 
-# I acknowledge that the nhl api exists and is a usable R package, however.. #
-# ..isolating specific statistics by teams and players during a 10 year period using the API was too difficult #
+#### Save data ####
+# change raw_data to whatever name you assigned when you downloaded it.
+write_csv(wifi_data, "inputs/data/wifi_data.csv") 
+
+
+
+#### Download data 2####
+
+df = list_package_resources("6678e1a6-d25f-4dff-b2b7-aa8f042bc2eb") |>
+  filter(name == 
+           "2023-WardProfiles-2011-2021-CensusData") |>
+  get_resource()
[…]

### 2024-04-16 scripts/02-data_cleaning.R

@@ -1,129 +1,71 @@
 #### Preamble ####
-# Purpose: Isolates the columns of interest, sums the data and cleans up the names of columns.
-# Author: Jason Ngo
-# Data: Apr 16 2023
-# Contact: jason_ngo@live.com
+# Purpose: Cleans the raw Toronto free public wifi data and 25 wards data from opendatatoronto r package
+# Author: Bernice Bao
+# Date: 23 January 2024
+# Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Some knowledge of NHL statistics terminology
-
+# Pre-requisites: run "01-download_data.R" before executing this file
 
 #### Workspace setup ####
 library(tidyverse)
-library(here)
-library(readxl)
+library(janitor)
+library(dplyr)
 
-league_data <- read_excel(here::here("inputs/data/Summaryleague_1122.xlsx"))
-
-## The datasets from nhl.com come in pages, so I will load them and merge them immediately ##
-shooting_p1 <- read_excel(here::here("inputs/data/shooting_1122p1.xlsx"))
-shooting_p2 <- read_excel(here::here("inputs/data/shooting_1122p2.xlsx"))
-shooting_p3 <- read_excel(here::here("inputs/data/shooting_1122p3.xlsx"))
-shooting_p4 <- read_excel(here::here("inputs/data/shooting_1122p4.xlsx"))
-master_shooting <- rbind(shooting_p1, shooting_p2) |> rbind(shooting_p3) |> rbind(shooting_p4)
-
-forwards_p1 <- read_excel(here::here("inputs/data/forwards_1122p1.xlsx"))
-forwards_p2 <- read_excel(here::here("inputs/data/forwards_1122p2.xlsx"))
-forwards_p3 <- read_excel(here::here("inputs/data/forwards_1122p3.xlsx"))
-master_forwards <- rbind(forwards_p1, forwards_p2) |> rbind(forwards_p3)
-
-dmen_p1 <- read_excel(here::here("inputs/data/dmen_1122p1.xlsx"))
-dmen_p2 <- read_excel(here::here("inputs/data/dmen_1122p2.xlsx"))
-dmen_p3 <- read_excel(here::here("inputs/data/dmen_1122p3.xlsx"))
-master_dmen <- rbind(dmen_p1, dmen_p2) |> rbind(dmen_p3)
 
 #### Clean data ####
+# load in raw data
[…]

### 2024-04-16 scripts/03-test_data.R

@@ -1,52 +1,20 @@
 #### Preamble ####
-# Purpose: Tests the master dataframes against bound checks defined in the simulation script
-# Author: Jason Ngo
-# Data: Apr 16 2023
-# Contact: jason_ngo@live.com
+# Purpose: Tests if the simulated data is within the required bounds
+# Author: Bernice Bao
+# Date: 24 January 2023 
+# Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Some knowledge of NHL statistics terminology
+# Pre-requisites: run 00-simulate_data.R
 
 #### Workspace setup ####
 library(tidyverse)
-library(here)
-
-league_data <- read.csv(here::here("outputs/data/league_data.csv"))
-master_dmen <- read.csv(here::here("outputs/data/master_dmen.csv"))
-master_forwards <- read.csv(here::here("outputs/data/master_forwards.csv"))
 
 #### Test data ####
 
-#### League dataset tests ####
+data = read_csv("outputs/data/simulated_data.csv")
 
-## Let's test if there are 12 columns, aka 12 seasons, in the league dataset ##
-league_data |> unique() |> count() == 12
+# Test1
+all(data$ward_number >= 1 & data$ward_number <= 25)
 
-## Let's test if average goals is less than 5 and greater than 1. AKA is it a realistic value ##
-league_data$Goals_Per_Game < 5
-league_data$Goals_Per_Game > 1
-
-## Let's test if average shots per game is less than 45 and greater than 20. AKA is it a realistic value ##
-league_data$Shots_Against > 20
-league_data$Shots_Against < 45
-
-## Let's test if average PP percentage is less than 40 and greataer than 5. AKA is it a realistic value ##
-league_data$PowerPlay_Pct > 5
-league_data$PowerPlay_Pct < 40
-
-## Let's test if mean shooting percentage is less than 20 and greataer than 5. AKA is it a realistic value ##
-league_data$Mean_Sh_Pct > 5
-league_data$Mean_Sh_Pct < 20
-
-## Let's test if average powerplay opportunities/game is less than 6 and greater than 1. AKA is it a realistic value ##
-league_data$PowerPlay_Game > 1
-league_data$PowerPlay_Game < 6
-
-#### Skater Dataset Tests #### 
-
-## Let's test if forward points per game is less than 2 and greater than 1
-master_forwards$Points_Per_Game > 1
[…]

### 2024-04-16 scripts/04-model.R

@@ -31,5 +31,3 @@ saveRDS(
   goals_model,
   file = here::here("outputs/models/goals_model.rds")
 )
-
-

### 2024-04-16 scripts/00-simulate_data.R
a

@@ -42,3 +42,4 @@ head(simulated_data)
 
 #### Save data ####
 write_csv(simulated_data,"outputs/data/simulated_data.csv")
+

### 2024-04-16 scripts/04-model.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Generates a linear model based on NHL data
-# Author: Jason Ngo
-# Data: Apr 16 2023
-# Contact: jason_ngo@live.com
+# Author: Bernice Bao
+# Data: Apr 16 2024
+# Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
 
 
@@ -11,9 +11,10 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-league_data <- read.csv(here::here("outputs/data/league_data.csv"))
+cleaned_ward_data <- read.csv(here::here("outputs/data/cleaned_ward_data.csv"))
+cleaned_wifi_data <- read.csv(here::here("outputs/data/cleaned_wifid_data.csv"))
 
-### Model data ####
+### Model data 1 ####
 goals_model <-
   stan_glm(
     formula = Goals_PG ~ SV_Pct + GA_Avg + Sh_Pct + PP_PG * PP_Pct,

### 2024-04-17 outputs/paper/paper.qmd
revise the formation of Preamble part

@@
 
-This research is divided into three main parts: `Introduction`(@sec-intro), `Data`(@sec-data), and `Conclusion`(@sec-conclusion).
+This research is divided into three main parts: `Introduction`(@sec-intro), `Data`(@sec-data), `Model` (@sec-model), `Results` (@sec-results), and `Discussion` (@sec-disc).
 Commencing with the Data section, a meticulous exposition ensues, delineating the origin of datasets garnered from the OpenDataToronto Library and expounding upon the procedural rigors applied for data refinement and analysis.
@@
 
-# Conclusion {#sec-conclusion}
+# Model {#sec-model}
+
+# Results {#sec-results}
 
@@
 
+# Discussion {#sec-disc}
+
 All in all, a significant number of people are increasingly relying on public Wi-Fi when they're out and about, underscoring their need for instant, hassle-free connectivity.
@@
 
+# Appendix 
+\newpage
+
 # References

### 2024-04-17 scripts/00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Get data on Location and description data of all City of Toronto free public WiFi locations and make table
 # Author: Bernice Bao
-# Date: 23 January 2024 
+# Date: 15 April 2024 
 # Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: NA

### 2024-04-17 scripts/01-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from the opendatatoronto r package
 # Author: Bernice Bao
-# Date: 23 January 2024
+# Date: Apr 16 2024
 # Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: NA

### 2024-04-17 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw Toronto free public wifi data and 25 wards data from opendatatoronto r package
 # Author: Bernice Bao
-# Date: 23 January 2024
+# Date: Apr 16 2024
 # Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: run "01-download_data.R" before executing this file

### 2024-04-17 scripts/03-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests if the simulated data is within the required bounds
 # Author: Bernice Bao
-# Date: 24 January 2023 
+# Date: Apr 16 2024
 # Contact: bernice.bao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: run 00-simulate_data.R
@@ -16,5 +16,7 @@ data = read_csv("outputs/data/simulated_data.csv")
 # Test1
 all(data$ward_number >= 1 & data$ward_number <= 25)
 
-#Test2
+# Test2
 all(data$ward %in% c("Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"))
+
+# Test3

### 2024-04-17 outputs/paper/paper.qmd
create new date frame for final usage

@@
 In this section, I will explore the diverse modes of `R`-based data collection and generation within the `RStudio` environment [@citeR; @citeRStudio].
-The data under scrutiny in this paper originates from the `opendatatoronto` R package [@citeODT], specifically drawing from the Toronto fire incidents dataset.
-The analysis benefits from additional functionalities harnessed from various R packages such as `tidyverse` [@citeTidyverse], `here` [@citeHere], `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], and `dplyr` [@citeDplyr].
+The data under scrutiny in this paper originates from the `opendatatoronto` R package [@citeODT], specifically drawing from theLocation and description data of Toronto free public WiFi dataset, as well as the 2021 Ward Profiles based on the 25-Ward model dataset.
+The analysis benefits from additional functionalities harnessed from various R packages such as `tidyverse` [@citeTidyverse] installed to gain access to other important R packages, `here` [@citeHere] created a path to specific saved files, `readr` [@citeReadr] read and imported data, `ggplot2` [@citeGgplot2] made the data visualizations, `knitr` [@citeKnitr] and `dplyr` [@citeDplyr] manipulated and cleaned data, and `modelsummary` [@citemodelSummary] to create summary tables.
 Further insights into the deployment of these packages will be expounded upon in the ensuing subsections.
 
-## Location and description data of Toronto free public WiFi
+## First Dataset: Location and description data of Toronto free public WiFi
 
@@
 
-## 2021 Ward Profiles based on the 25-Ward model
+## Second Dataset: 2021 Ward Profiles based on the 25-Ward model
 
@@
 York South-Weston (Ward 5) at $67,964, and Scarborough Centre (Ward 21) at $70,624.
+
+## Variables of Interest
 

### 2024-04-17 scripts/00-simulate_data.R

@@ -9,18 +9,11 @@
 
 #### Workspace setup ####
 library(tidyverse)
-install.packages("opendatatoronto")
-install.packages("knitr")
-
 library(knitr)
 library(janitor)
 library(lubridate)
 library(opendatatoronto)
 library(tidyverse)
-
-citation() # Get the citation information for R
-citation("ggplot2") # Get citation information for a package
-
 
 #### Simulate data ####
 set.seed(85)

### 2024-04-17 scripts/02-data_cleaning.R

@@ -52,11 +52,13 @@ cleaned_ward_data <-
 # move the variables to the correct position
 
 cleaned_ward_data <- data.frame(cleaned_ward_data)
-colnames(cleaned_ward_data ) <- cleaned_ward_data[1,]
+colnames(cleaned_ward_data) <- cleaned_ward_data[1,]
 cleaned_ward_data <- cleaned_ward_data[-1,]
 rownames(cleaned_ward_data) <- 1:25
 
-cleaned_ward_data <- cleaned_ward_data |> mutate(WardNumber = c(1:25))
+cleaned_ward_data <- cleaned_ward_data |> mutate(WardNumber = c(1:25)) 
+
+cleaned_ward_data$WardNumber <- as.character(cleaned_ward_data$WardNumber)
 
 # rename the variables
 cleaned_ward_data <-
@@ -65,7 +67,43 @@ cleaned_ward_data <-
 
 head(cleaned_ward_data)
 
+
 #### Save data ####
 # write cleaned data as csv
 write_csv(cleaned_ward_data, "outputs/data/cleaned_ward_data.csv")
+
+
+
+
+
+
+#### Create final data ####
+
+# Find out building type combination counted in each ward
+
+new_data <-
+  cleaned_wifi_data |> 
+  group_by(WardNumber) |> 
+  mutate(WardNumber = str_replace(WardNumber, "05", "5"))|> 
+  summarise(Numer_of_free_wifi = n()) |> 
+  arrange(WardNumber)
+  
+
+head(new_data)
+# Add new variables of building type combination counted in each ward
+
+final_data <- full_join(x= cleaned_ward_data, y= new_data, by="WardNumber")
+
+final_data <-
+  cleaned_wifi_data |>
+  filter(`City of Toronto Profiles` == "Total - Age" | `City of Toronto Profiles` == "Total - Household total income groups in 2020 for private households - 25% sample data")|> # filter on the population and income
+  select(`City of Toronto Profiles`, ...3, ...4, ...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17, ...18, ...19, ...20, ...21, ...22, ...23, ...24, ...25, ...26, ...27) |># only select the relevant columns
+  t()
+
+head(final_data)
+
+#### Save data ####
+# write cleaned data as csv
+write_csv(final_data, "outputs/data/final_data.csv")
+
 

### 2024-04-17 scripts/02-data_cleaning.R
repair a mistake about summarising the cleaned data

@@ -85,10 +85,11 @@ new_data <-
   cleaned_wifi_data |> 
   group_by(WardNumber) |> 
   mutate(WardNumber = str_replace(WardNumber, "05", "5"))|> 
-  summarise(Numer_of_free_wifi = n()) |> 
-  arrange(WardNumber)
+  summarise(Numer_of_free_wifi = n())|> 
   
+  arrange(WardNumber)
 
+  
 head(new_data)
 # Add new variables of building type combination counted in each ward
 

### 2024-04-17 scripts/02-data_cleaning.R
delete unnecessary steps

@@ -86,20 +86,14 @@ new_data <-
   group_by(WardNumber) |> 
   mutate(WardNumber = str_replace(WardNumber, "05", "5"))|> 
   summarise(Numer_of_free_wifi = n())|> 
-  
   arrange(WardNumber)
-
+  
   
 head(new_data)
 # Add new variables of building type combination counted in each ward
 
 final_data <- full_join(x= cleaned_ward_data, y= new_data, by="WardNumber")
 
-final_data <-
-  cleaned_wifi_data |>
-  filter(`City of Toronto Profiles` == "Total - Age" | `City of Toronto Profiles` == "Total - Household total income groups in 2020 for private households - 25% sample data")|> # filter on the population and income
-  select(`City of Toronto Profiles`, ...3, ...4, ...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17, ...18, ...19, ...20, ...21, ...22, ...23, ...24, ...25, ...26, ...27) |># only select the relevant columns
-  t()
 
 head(final_data)
 

### 2024-04-17 scripts/02-data_cleaning.R
modificate the linear regression model

@@ -85,7 +85,7 @@ new_data <-
   cleaned_wifi_data |> 
   group_by(WardNumber) |> 
   mutate(WardNumber = str_replace(WardNumber, "05", "5"))|> 
-  summarise(Numer_of_free_wifi = n())|> 
+  summarise(Number_of_free_wifi = n())|> 
   arrange(WardNumber)
   
   

### 2024-04-17 scripts/04-model.R

@@ -11,14 +11,13 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-cleaned_ward_data <- read.csv(here::here("outputs/data/cleaned_ward_data.csv"))
-cleaned_wifi_data <- read.csv(here::here("outputs/data/cleaned_wifid_data.csv"))
+final_data <- read.csv(here::here("outputs/data/final_data.csv"))
 
 ### Model data 1 ####
-goals_model <-
+wifi_model <-
   stan_glm(
-    formula = Goals_PG ~ SV_Pct + GA_Avg + Sh_Pct + PP_PG * PP_Pct,
-    data = league_data,
+    formula = Number_of_free_wifi ~ Population + Income,
+    data = final_data,
     family = gaussian(),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -29,6 +28,7 @@ goals_model <-
 
 #### Save model ####
 saveRDS(
-  goals_model,
-  file = here::here("outputs/models/goals_model.rds")
+  wifi_model,
+  file = here::here("outputs/models/wifi_model.rds")
 )
+

### 2024-04-17 scripts/04-model.R
add 2 more model for different variables

@@ -9,11 +9,13 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(modelsummary)
 
 #### Read data ####
 final_data <- read.csv(here::here("outputs/data/final_data.csv"))
 
-### Model data 1 ####
+
+#### Model data 1 ####
 wifi_model <-
   stan_glm(
     formula = Number_of_free_wifi ~ Population + Income,
@@ -22,13 +24,51 @@ wifi_model <-
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 1122
+    seed = 123
   )
 
 
+## creating simulated data for the model
+set.seed(853)
+
+
+#### Linear Model data 2 ####
+population_model <-
+  lm(
+    Number_of_free_wifi ~ Population,
+    data = final_data,
+  )
+
+summary(population_model)
+modelsummary(population_model)
+
+
+#### Linear Model data 3 ####
+income_model <-
+  lm(
+    Number_of_free_wifi ~ Population,
+    data = final_data,
+  )
+
+summary(income_model)
+modelsummary(income_model)
+
+
 #### Save model ####
+
 saveRDS(
   wifi_model,
   file = here::here("outputs/models/wifi_model.rds")
 )
 
+# LINEAR MODEL 2
+saveRDS(
+  population_model,
+  file = "outputs/models/population_model.rds"
+)
+
+# LINEAR MODEL 3
+saveRDS(
+  income_model,
+  file = "outputs/models/income_model.rds"
+)

### 2024-04-17 outputs/paper/paper.qmd
correct some mistakes and knit the latest paper

@@
   - Bernice Bao
-thanks: "Code and data are available at: <https://github.com/BerniceBao/Free-Public-WiFi-Distribution-in-Toronto>
- <https://open.toronto.ca/dataset/city-of-toronto-free-public-wifi/>, and <https://open.toronto.ca/dataset/ward-profiles-25-ward-model/>."
+thanks: "Code and data are available at: <https://github.com/BerniceBao/Free-Public-WiFi-Distribution-in-Toronto>."
 date: 15 April 2024
@@
 
+```{r}
+#| include: false
+
+# IMPORT MODEL FROM OUTPUTS FOLDER
+wifi_model <- readRDS(here::here("outputs/models/wifi_model.rds"))
+population_model <- readRDS(here::here("outputs/models/population_model.rds"))
+income_model <- readRDS(here::here("outputs/models/income_model.rds"))
+
+```
+
+The goal of my modelling strategy is twofold.
+Firstly, I want to analyze to what effect power play opportunities, power play effectiveness, save percentage, goals against average and shooting percentages all have on average goals scored per game.
+Secondly, I want to use these results to predict the average goals scored per game for the next season.
+
+In the section below, the Bayesian analysis model used to investigate the multiple linear regression model of average goals scored per game is detailed.
+
+## Model set-up
+
+Define $y_i$ as the leaguewide average goals scored per game.
+Then $\beta_1$ is the leaguewide average power plays per game, $\gamma_1$ is leaguewide powerplay percentage, $\theta_1$ is leaguewide save percentage, $\phi_1$ is the leaguewide goals against average, and $\kappa_1$ is the leaguewide average shooting percentage.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \theta_i + \phi_i + \kappa_i + \beta_i * \gamma_i \\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\theta &\sim \mbox{Normal}(0, 2.5) \\
+\phi &\sim \mbox{Normal}(0, 2.5) \\
+\kappa &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in `R` [@citeR] using the `rstanarm` package of [@citestan]. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between average goals scored per game and power play opportunities, power play percentage, goals against average, and shooting percentage.
+The higher those four variables are, the more goals there should be.
+In particular, there should also be a negative relationship between goals scored and save percentage.
+With a lower save percentage, there are less saves that are occuring which means there should be more goals.
+Additionally, the two power play variables are interacted with eachother, since they are directly related to one another.
+There cannot be a power play percentage without a power play opportunity.
+
 # Results {#sec-results}
@@
 #| label: fig-1
-#| fig-cap: Relationship between Population of each Ward and Amount of Free Public WiFi Location
+#| fig-cap: Linear regression model of the number of produced ASL words by the range of children's ages during their early developmental years.
 #| echo: false
 #| warning: false
-#| message: false
 
-analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))
-
-analysis_data_1 |> 
-  ggplot(aes(x = Population, y = WardNumber)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Population of each Ward",
-       y = "Ward Number")
+population_model |>
+  ggplot(aes(x = Population, y = Number_of_free_wifi)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(
+    method = "lm",
+    se = TRUE,
+    color = "black",
+    linetype = "dashed",
+    formula = "y ~ x"
+  ) +
+  labs(
+    x = "Average Income of a Ward",
+    y = "Number of Free Wifi in a ward"
+  ) +
+  theme_classic()
 
@@
 
-From this graph, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.
+Model 1, as illustrated in @fig-prodwordsmodel, displays the correlation between a child's age in months and the number of ASL words they can produce at a given age.
+The scatterplot depicts the general trend of vocabulary growth as a child grows in age, further emphasized by the regression line.
+The model indicates that as the age of a child increases, the number of vocabulary items that that child produces increases as well.
+The regression line allows for the prediction of future outcomes based on the known data available from the dataset.
+By utilizing this model, we can estimate when a child will be able to produce a certain number of ASL vocabulary words or phrases at a certain age.
+
+From @fig-1, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.
 
@@
 #| label: fig-2
-#| fig-cap: Relationship between Income of each Ward and Amount of Free Public WiFi Location
+#| fig-cap: Linear regression model of the number of produced ASL words by the range of children's ages during their early developmental years.
 #| echo: false
 #| warning: false
-#| message: false
 
-analysis_data_1 <- read_csv(here::here("outputs/data/cleaned_ward_data.csv"))
-
-analysis_data_1 |> 
-  ggplot(aes(x = Income, y = WardNumber)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Income of each Ward",
-       y = "Ward Number")
+income_model |>
+  ggplot(aes(x = Income, y = Number_of_free_wifi)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(
+    method = "lm",
+    se = TRUE,
+    color = "black",
+    linetype = "dashed",
+    formula = "y ~ x"
+  ) +
+  labs(
+    x = "Average Income of a Ward",
+    y = "Number of Free Wifi in a ward"
+  ) +
+  theme_classic()
 
 ```
-
-From this graph, we can clearly see that Spadina-Fort York (Ward 10) has the greatest average income, Toronto Centre (Ward 13) has the second greatest average income, and Etobicoke-Lakeshore (Ward 3) has the third greatest average income.
+From @fig-2, we can clearly see that Spadina-Fort York (Ward 10) has the greatest average income, Toronto Centre (Ward 13) has the second greatest average income, and Etobicoke-Lakeshore (Ward 3) has the third greatest average income.
 

### 2024-04-17 scripts/04-model.R

@@ -46,7 +46,7 @@ modelsummary(population_model)
 #### Linear Model data 3 ####
 income_model <-
   lm(
-    Number_of_free_wifi ~ Population,
+    Number_of_free_wifi ~ Income,
     data = final_data,
   )
 
@@ -72,3 +72,4 @@ saveRDS(
   income_model,
   file = "outputs/models/income_model.rds"
 )
+

### 2024-04-17 outputs/paper/paper.qmd
Add more details about data parts

@@
 #### Read in the data ####
+cleaned_ward_data <-
+  read_csv(
+    file = here("outputs/data/cleaned_ward_data.csv"),
+    show_col_types = FALSE
+  )
+
 cleaned_wifi_data <-
@@
   )
-#### Read in the data ####
-cleaned_ward_data <-
+
+final_data <-
   read_csv(
-    file = here("outputs/data/cleaned_ward_data.csv"),
+    file = here("outputs/data/final_data.csv"),
+    show_col_types = FALSE
+  )
+
+building_data <-
+  read_csv(
+    file = here("outputs/data/building_data.csv"),
     show_col_types = FALSE
@@
 
+```{r}
+#| label: fig-interest
+#| fig-cap: The number of free Wifi available in different wards in Toronto
+#| echo: false
+#| warning: false
+
+library(tidyverse)
+
+final_data |> ggplot(aes(x = WardNumber, y = Number_of_free_wifi)) +
+  geom_col() +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
+  labs(
+    title = "The number of free Wifi available in different wards in Toronto",
+    x = "Ward Number",
+    y = "Number of Free Public Wifi"
+  )
+
+```
+
 # Model {#sec-model}
@@
 
-## Demand
+## Building Type
 
 **Societal need for public Wi-Fi differs according to building types**
+
+While I talked about the increase of the amount of productive forwards and defensemen, here I will discuss the increase in the amount of good forwards and compare them to the amount of elite forwards.
+To be classified as an elite forward, usually they would need to finish the season at around a 1.4 PPG pace.
+Because of how hard of a feat it is to achieve a sustained 1.4 PPG pace, there are very few elite forwards in the NHL.
+@fig-building shows the amount of forwards that finish in multiple PPG ranges.
+
+```{r}
+#| label: fig-building
+#| fig-cap: The number of free Wifi in Toronto by building type
+#| echo: false
+#| warning: false
+
+library(tidyverse)
+
+building_data |> 
+  ggplot(aes(x = BuildingType, y = Number_of_free_wifi)) +
+  geom_col() +
+  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
+  labs(
+    title = "The number of free Wifi in Toronto by building type",
+    x = "Building Types",
+    y = "Number of Free Public Wifi"
+  )
+
+```
+
+As evidenced, the amount of elite players have remained consistent over the seasons, although the big spike in the 2021-22 season may be a sign that times are changing.
+The bigger story here though, is the amount of players that finish in the 1.2 and 1.4 range from 2017-18 beyond.
+There is a dramatic spike in these productive players from before 2017-18 to that season and beyond.
+The graphs here show that there while the ceiling of player talent has remained the same, the floor is steadily increasing.
 
@@
 
-Model 1, as illustrated in @fig-prodwordsmodel, displays the correlation between a child's age in months and the number of ASL words they can produce at a given age.
+Model 1, as illustrated in @fig-1, displays the correlation between a child's age in months and the number of ASL words they can produce at a given age.
 The scatterplot depicts the general trend of vocabulary growth as a child grows in age, further emphasized by the regression line.

### 2024-04-17 scripts/00-simulate_data.R

@@ -34,5 +34,5 @@ simulated_data <-
 head(simulated_data)
 
 #### Save data ####
-write_csv(simulated_data,"outputs/data/simulated_data.csv")
+write_csv(simulated_data,"inputs/data/simulated_data.csv")
 

### 2024-04-17 scripts/02-data_cleaning.R

@@ -74,12 +74,38 @@ write_csv(cleaned_ward_data, "outputs/data/cleaned_ward_data.csv")
 
 
 
+#### Create building data ####
 
+building_data <-
+  cleaned_wifi_data |> 
+  group_by(BuildingType) |> 
+  summarise(Number_of_free_wifi = n())|> 
+  arrange(desc(Number_of_free_wifi)) |>
+  mutate(BuildingType = str_replace(BuildingType, "Arena,Artificial Ice Rink Building,Community Recreation Centre,Outdoor Pool Building,Washroom Building", "A,AIRB,CRC,OPB,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Arena,Artificial Ice Rink Building,Washroom Building", "A,AIRB,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Arena,Washroom Building", "A,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Artificial Ice Rink Building,Changeroom Building,Community Recreation Centre,Washroom Building", "AIRB,CB,CRC,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Artificial Ice Rink Building,Community Recreation Centre,Washroom Building", "AIRB,CRC,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Community Recreation Centre,Outdoor Pool Building,Washroom Building", "CRC,OPB,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Community Recreation Centre,School", "CRC,School")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Artificial Ice Rink Building,Community Recreation Centre,Outdoor Pool Building,Washroom Building", "AIRB,CRC,OPB,WB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Arena,Community Recreation Centre", "Arena,CRC")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Arena,Community Recreation Centre,Outdoor Pool Building", "Arena,CRC,OPB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Community Recreation Centre,Outdoor Pool Building", "CRC,OPB")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Community Recreation Centre", "CRC")) |>
+  mutate(BuildingType = str_replace(BuildingType, "Outdoor Pool Building", "OPB")) |>
[…]

### 2024-04-17 scripts/04-model.R

@@ -14,7 +14,6 @@ library(modelsummary)
 #### Read data ####
 final_data <- read.csv(here::here("outputs/data/final_data.csv"))
 
-
 #### Model data 1 ####
 wifi_model <-
   stan_glm(
@@ -53,7 +52,6 @@ income_model <-
 summary(income_model)
 modelsummary(income_model)
 
-
 #### Save model ####
 
 saveRDS(
@@ -72,4 +70,5 @@ saveRDS(
   income_model,
   file = "outputs/models/income_model.rds"
 )
+
 

### 2024-04-17 outputs/paper/paper.qmd
Add more detailed explanations for the model parts

@@
 
-### Model justification
+## Model justification
 
@@
   labs(
-    title = "The number of free Wifi in Toronto by building type",
     x = "Building Types",
@@
 #| label: fig-1
-#| fig-cap: Linear regression model of the number of produced ASL words by the range of children's ages during their early developmental years.
+#| fig-cap: Linear regression model of the number of Free public wifi by the populations of all wards in Toronto.
 #| echo: false
@@
 #| label: fig-2
-#| fig-cap: Linear regression model of the number of produced ASL words by the range of children's ages during their early developmental years.
+#| fig-cap: Linear regression model of the number of Free public wifi by the Average Income of all wards in Toronto.
 #| echo: false

### 2024-04-17 outputs/paper/paper.qmd
Add details about explanation of model

@@
 #### Workspace set-up ####
+tinytex::install_tinytex(force = TRUE)
 library(tidyverse)
@@
 library(here)
+library(modelsummary)
 
@@
 
-In this section, I will explore the diverse modes of `R`-based data collection and generation within the `RStudio` environment [@citeR; @citeRStudio].
-The data under scrutiny in this paper originates from the `opendatatoronto` R package [@citeODT], specifically drawing from theLocation and description data of Toronto free public WiFi dataset, as well as the 2021 Ward Profiles based on the 25-Ward model dataset.
+In this section, I'll explore the R-based data collection process and generate the visualizations in the RStudio environment [@citeR;@citeRStudio].
+All the datasets studied in this paper come from "Open Data Toronto" [@citeODT], a data source allowing researchers and policymakers to access and utilize the data which can be downloaded for free from its website and GitHub repository, in order to finish various statistical analyses, research studies, and decision-making processes.
+The two main datasets used in this paper are @toronto_wifi_dataset, which provides descriptive data such as the location and building type of Toronto free public wifi, and @toronto_ward_profiles, which contains economic, population and other social information of all 25 wards in Toronto collected in 2021.
+
 The analysis benefits from additional functionalities harnessed from various R packages such as `tidyverse` [@citeTidyverse] installed to gain access to other important R packages, `here` [@citeHere] created a path to specific saved files, `readr` [@citeReadr] read and imported data, `ggplot2` [@citeGgplot2] made the data visualizations, `knitr` [@citeKnitr] and `dplyr` [@citeDplyr] manipulated and cleaned data, and `modelsummary` [@citemodelSummary] to create summary tables.
@@
 
-The data I mainly used to generate the visualizations in this paper is City of Toronto Free Public WiFi from Open Data Toronto [@citeODT] that is an open source tool designed to be publicly available and encourage the development of valuable insights.
-The dataset, released by Toronto's Department of Information & Technology, captures information on location and description data free public WiFi locations in the City of Toronto and is updated monthly.
+The data I primarily use to generate visualizations in this article is an open source tool designed to be publicly available and to encourage the development of valuable insights.
+The dataset[@toronto_wifi_dataset], published by Toronto's Department of Information Technology, collects location information and descriptive data for free public WiFi locations in the City of Toronto and is updated monthly, with the latest update on April 15, 2024.
+According to the data characteristics described on the portal, these data include address, building type, building name, public name, zip code, whether there is public wifi, owner, ward, ward number, and other geographical information.
 
-The data used for analysis in this article is as of January 3, 2024.
-According to the data characteristics described on the portal, this data should include information such as address, building type, building name, postal code, whether there is public wifi, etc.
-Through systematic analysis, it is found that only building type is the main subject variable that can be used to study the influencing factors of distribution.
-In order to quickly extract available data, only columns with reasonable data are retained during the data cleaning process (see Table 1).
-Basic data cleaning of column values to shorten the description of different types of clinics and improve readability.
-I create a statistic on the basis of monthly groups, using kable() from knitr to create @tbl-wifi.
+Through further systematic analysis, among the statistical data about public wifi, our analysis is particularly interested in building type, ward and ward number, which are very reasonable as the main subject variables that can be used to study the influencing factors of public wifi distribution.
+Among them, ward and ward number belong to the nature of the distribution of public wifi in Toronto itself, and do not take into account the influencing factors to be explored, but provide a new idea that can be connected with more social information associated with ward to explore more influential factors that can be studied.
+
+Since I only focused on the impact of building type and public wifi layout in Toronto, I kept only the building type, ward, and ward number columns during the data cleaning process in order to quickly extract available data.
+In addition, I carried out basic data cleaning on column values and deleted building data and NA information that did not contain wifi to improve readability, then summaried the building types and used the able() of knitr to create @tbl-building:
 
 ```{r}
-#| label: tbl-wifi
-#| tbl-cap: Sample of cleaned Free Public WiFi data
+#| label: tbl-building
+#| tbl-cap: Sample of building type data
 #| echo: false
 
-cleaned_wifi_data %>%
+building_data %>%
   head(10) %>%
   kable(
-    col.names = c("Building type", "Ward", "Ward Number"),
+    col.names = c("Building Type", "Number of Free Wifi"),
     digits = 1
@@
 
-From the Location and description data of Toronto free public WiFi, as of January 3, 2024, there are 194 free public WiFi location in the City of Toronto.
-Further breakdown shows that there are 27 free public WiFi location that are set in Arena, 74 free public WiFi location that are set in Community Recreation Centre, 17 free public WiFi location that are set in Outdoor Pool Building clinics, and  100 free public WiFi location that are set in library.
+From @tbl-final, there are 198 free public WiFi location in the City of Toronto in total.
+Further breakdown shows that there are  100 free public WiFi location that are set in library, 74 free public WiFi location that are set in Community Recreation Centre(CRC), 27 free public WiFi location that are set in Arena, 17 free public WiFi location that are set in Outdoor Pool Building (OPB).
 Since the same wifi may cover multiple buildings, the sum of the number of public wifi corresponding to each building type is greater than the total number of public wifi in Toronto.
@@
 Besides, Community Recreation Centre has the second most amount of free public WiFi.
-This information is indicative of the efforts or infrastructure investments made by Community Recreation Centres to offer accessible and free internet connectivity to the public.
+This information is indicative of the efforts or infrastructure investments made by Toronto goverment to offer accessible and free internet connectivity to the public.
 
@@
 
-In order to deeply understand the factors affecting the distribution of public wifi in Toronto, I will study the population and income of each ward in Toronto on this basis.
-Therefore, the regional profile dataset (Data 2021), based on the 2021 Census data, was also included in the analysis.
-The dataset, published by City Planning, was last updated on January 3, 2024.
-The dataset contains demographic, social and economic information such as income and population for each ward.[@toronto-ward-profiles]
+In addition to building types, I also studied the population and income of each ward in Toronto as another two variables to deeply understand the factors affecting the distribution of public wifi in Toronto.
+Therefore, the ward profile dataset [@toronto-ward-profiles] published by City Planning based on the 2021 Census data was also included in the analysis.
+This dataset contains demographic, social and economic information such as income and population for each ward, and last updated on January 3, 2024.
 
-Ward profile data is stored in an Excel file with multiple tabs.
-The relevant data used for this analysis is included as a variable in the first TAB, the 2021 Census.
-Therefore, only the data from this TAB is downloaded for analysis.
-Further data cleansing was done to transform the data, retaining only information related to income and population levels in each ward (see @tbl-ward).
-Based on the Toronto Open Data Portal, the package has a CSV file, 25 ward names and numbers, which contains the mapping between ward codes and ward names.
-I create statistic to visualize these data, using kable() from knitr to create @tbl-ward.
+Ward profile data is stored in an Excel file with multiple tabs, and the relevant data used for this analysis is included as a variable in the first TAB, the 2021 Census.
+Data included population, age, race, language and median household income.
+Median household income is the pre-tax income a family receives from employment, investment and government support.
+Median income is used because it is less sensitive to very high and very low income values than the average.
+
+This research is concerned with the impact of population and income on the distribution of free wifi, so I only choose to analyze these two variables as my research objects, and created statistic to visualize the data.
+we were able to get initial valid information but still need more clarity for further data analysis, so I established a new variable, "number of free wifi", to quantify the performance of free wifi in different building types.
+Then I used kable() from knitr to get @tbl-final:
 
 ```{r}
-#| label: tbl-ward
-#| tbl-cap: Sample of cleaned Ward of Toronto data
+#| label: tbl-final
+#| tbl-cap: Sample of final ward data
 #| echo: false
 
-cleaned_ward_data %>%
+final_data %>%
   head(10) %>%
   kable(
-    col.names = c("Population", "Income", "Ward Number"),
+    col.names = c("Population", "Income", "Ward Number", "Number of Free Wifi"),
     digits = 1
@@
 
-There are 25 wards in the City of Toronto. Based 2021 census data, the wards with the highest population are:
-Etobicoke-Lakeshore (Ward 3) at 139920, Spadina-Fort York (Ward 10) at 135400, and Etobicoke Centre (Ward 2) at 117200. The wards with the lowest population are: Scarborough North (Ward
-23) at 94025, Don Valley East (Ward 16) at 94335, and Don Valley West (Ward
-15) at 101025.
+From @tbl-final, there are 25 wards in total in the City of Toronto, and the wards with the highest population are: Etobicoke-Lakeshore (Ward 3) at 139920, Spadina-Fort York (Ward 10) at 135400.
+The wards with the lowest population are: Scarborough North (Ward 23) at 94025, Don Valley East (Ward 16) at 94335.
+The wards with the highest household income are: Spadina-Fort York (Ward 10) at $80730$, Toronto Centre (Ward 13) at $68965$.
+The wards with the lowest household income are: Humber River-Black Creek (Ward 7) at $65,458$, York South-Weston (Ward 5) at $67,964$.
 
-The wards with the highest household income are: Spadina-Fort York (Ward 10) at $ 80730,
-Toronto Centre (Ward 13) at $68965, and Etobicoke-Lakeshore (Ward 3) at $65575. The
-wards with the lowest household income are: Humber River-Black Creek (Ward 7) at $65,458,
-York South-Weston (Ward 5) at $67,964, and Scarborough Centre (Ward 21) at $70,624.
-
-## Variables of Interest
+It intuitively shows the social developments of 25 neighbourhoods in Toronto, including population, median income, and number of wifi, providing data support for the following study of the linear relationship between the three.
+It is important to note that, according to the database, more populated neighborhoods do not necessarily have more median income, which laid the foundation for our decision to separate the two variables to analyze the impact of public wifi distribution.
 
@@
   ) 
-
 ```
+@fig-interest shows the data visualization shows that the ward1, 3, and 7 have significantly more than the median number of public wifi, 12, 13, and 14 respectively, which is more than 1/3 times more than the average of 8.
+From this graph, we can observe that except for the ward 1, 3 and 7, the number of public wifi in other wards is basically in the same range, and there is no particularly serious shortage.
+This indicates that the level of public Internet accessibility strength is generally consistent across wards in Toronto.
 
@@
 
-As evidenced, the amount of elite players have remained consistent over the seasons, although the big spike in the 2021-22 season may be a sign that times are changing.
-The bigger story here though, is the amount of players that finish in the 1.2 and 1.4 range from 2017-18 beyond.
-There is a dramatic spike in these productive players from before 2017-18 to that season and beyond.
-The graphs here show that there while the ceiling of player talent has remained the same, the floor is steadily increasing.
+Library has the highest number of free public WiFi, which means that, in comparison to other types of buildings in Toronto, libraries offer the highest availability or coverage of free public WiFi services.
+This suggests that libraries prioritize or invest in providing accessible and free internet connectivity to the public within Toronto.
+Besides, Community Recreation Centre has the second most amount of free public WiFi.
+This information is indicative of the efforts or infrastructure investments made by Community Recreation Centres to offer accessible and free internet connectivity to the public.
 
@@
   labs(
-    x = "Average Income of a Ward",
+    x = "Population of a Ward",
     y = "Number of Free Wifi in a ward"
@@
 ```
+The first linear regression model, as illustrated in @fig-1, explores the relationship between the number of free public WiFi locations and the population of wards across Toronto.
+The scatterplot depicts the general trend of increased number of public wifi as bigger population, further emphasized by the regression line.
+The model indicates that as the population of a ward gets bigger, the number of free public wifi increases as well.
 
-Model 1, as illustrated in @fig-1, displays the correlation between a child's age in months and the number of ASL words they can produce at a given age.
-The scatterplot depicts the general trend of vocabulary growth as a child grows in age, further emphasized by the regression line.
-The model indicates that as the age of a child increases, the number of vocabulary items that that child produces increases as well.
-The regression line allows for the prediction of future outcomes based on the known data available from the dataset.
-By utilizing this model, we can estimate when a child will be able to produce a certain number of ASL vocabulary words or phrases at a certain age.
+However, the coefficient for population (0.00) indicates that there is no statistically significant linear relationship between the population of a ward and the number of free public WiFi locations it hosts, as the coefficient's p-value exceeds conventional thresholds for significance.
+This finding suggests that the population size alone may not be a strong predictor of the availability of free public WiFi in Toronto's wards.
+The model's low R-squared value of 0.045 and adjusted R-squared value of 0.003 indicate that only a very small proportion of the variability in the number of free public WiFi locations can be explained by the population size of the wards.
+The model's RMSE (Root Mean Square Error) of 2.26 suggests that the model's predictions deviate from the actual values by an average of 2.26 free public WiFi locations.
+
+In summary, while the model attempts to assess the relationship between population and the availability of free public WiFi, its findings indicate that population alone is not a strong predictor of the number of free public WiFi locations in Toronto's wards.
 
 From @fig-1, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-model1
+#| tbl-cap: "A summary table of the first linear regression model"
+
+modelsummary::modelsummary(
+  list("Population of a Ward" = population_model),
+  fmt = 2
+)
+```
+
+@tbl-model1 displays the coefficients of the predictor variables of the linear regression model.
+All the variables included in the model have p-values less than 0.05, indicating that they have a significant impact on the permanent immigrant inflows rate.
+Therefore, all the variables are retained in the final model.
+For every 1% increase in the Employment Rate, there is a 0.04% increase in the Permanent Inflows Rate.
+The relationship between Employment Rate and Permanent Inflows Rate is positive, which means that as the Employment Rate increases, the Permanent Inflows Rate also tends to increase.
+The slope coefficient of 0.04 indicates that for every 1% increase in the Employment Rate, we can expect a 0.04% increase in the Permanent Inflows Rate, all else being equal.
+As the "Year" variable increases by 1, the estimated permanent immigrant inflow increases by 0.04% on average, holding other predictor variables constant.
+However, it is important to note that this is a correlation and does not necessarily imply causation.
+Other factors, such as government policies and global economic conditions, may also have an impact on Permanent Inflows Rate.
 
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-model2
+#| tbl-cap: "A summary table of the second linear regression model"
+
+modelsummary::modelsummary(
+  list("Average Income of a Ward" = income_model),
+  fmt = 2
+)
+```
+
+@tbl-model2 displays the coefficients of the predictor variables of the linear regression model.
+All the variables included in the model have p-values less than 0.05, indicating that they have a significant impact on the permanent immigrant inflows rate.
+Therefore, all the variables are retained in the final model.
+For every 1% increase in the Employment Rate, there is a 0.04% increase in the Permanent Inflows Rate.
+The relationship between Employment Rate and Permanent Inflows Rate is positive, which means that as the Employment Rate increases, the Permanent Inflows Rate also tends to increase.
+The slope coefficient of 0.04 indicates that for every 1% increase in the Employment Rate, we can expect a 0.04% increase in the Permanent Inflows Rate, all else being equal.
+As the "Year" variable increases by 1, the estimated permanent immigrant inflow increases by 0.04% on average, holding other predictor variables constant.
+However, it is important to note that this is a correlation and does not necessarily imply causation.
+Other factors, such as government policies and global economic conditions, may also have an impact on Permanent Inflows Rate.
+
 We can conclude that there is no observable correlation between the number of Toronto public WiFi and the average household income in a ward.

### 2024-04-18 outputs/paper/paper.qmd
Finish most of the work

@@
 date-format: long
-abstract: "In the era of increasing digital dependence, understanding the intricacies of urban internet connectivity is pivotal for fostering an effective and technologically equitable society.
-Public WiFi is a vital tool in modern society, offering internet access and supporting economic development, education, and social connections.
-In response to the increasing importance of internet connectivity in contemporary urban life, I investigate the distribution of free public WiFi in Toronto.
-This paper aims to provide key statistics on the distribution of public WiFi in Toronto and analyze the impact of building types, demographics, and economic influences on its distribution to gain as comprehensive a picture as possible of the role and impact of public WiFi in our lives."
+abstract: "Public WiFi is a vital tool in modern society, offering no-cost internet access and supporting economic, social and educational development.
+In response to the increasing importance of public free wifi in contemporary urban life, this paper investigates the distribution and influencing factors of free public WiFi in Toronto.
+Drawing on datasets from Open Data Toronto, the paper focuses on the association between the building types of the place with public wifi, population and income of each ward in Toronto and number of  Toronto public free wifi by creating linear regression models, aimed to shed light on the correlation between free public WiFi distribution and building types, population, and income, and put forward relevent suggestions for the future development.
+Understanding these findings is crucial for policymakers, urban planners, and community advocates developing digital public Internet and foster a more equitable society."
 format: pdf
@@
 
-The data I primarily use to generate visualizations in this article is an open source tool designed to be publicly available and to encourage the development of valuable insights.
-The dataset[@toronto_wifi_dataset], published by Toronto's Department of Information Technology, collects location information and descriptive data for free public WiFi locations in the City of Toronto and is updated monthly, with the latest update on April 15, 2024.
-According to the data characteristics described on the portal, these data include address, building type, building name, public name, zip code, whether there is public wifi, owner, ward, ward number, and other geographical information.
+@toronto_wifi_dataset, the dataset I primarily use to generate visualizations is an open source tool designed to be publicly available and to encourage the development of valuable insights.
+It published by Toronto's Department of Information Technology, collects location information and descriptive data for free public WiFi locations in the City of Toronto and is updated monthly, with the latest update on April 15, 2024.
+These data include address, building type, building name, public name, zip code, whether there is public wifi, owner, ward, ward number, and other geographical information.
 
-Through further systematic analysis, among the statistical data about public wifi, our analysis is particularly interested in building type, ward and ward number, which are very reasonable as the main subject variables that can be used to study the influencing factors of public wifi distribution.
-Among them, ward and ward number belong to the nature of the distribution of public wifi in Toronto itself, and do not take into account the influencing factors to be explored, but provide a new idea that can be connected with more social information associated with ward to explore more influential factors that can be studied.
+Among all statistical data about free public wifi, our analysis is particularly interested in "building type", "ward" and "ward number", which are considered as the main subject variables of our study.
+Analyzing public WiFi distribution by building type allows us to understand where internet access is concentrated and which locations may be underserved.
+This insight can inform decisions about where to expand infrastructure, and how to ensure equitable access to digital resources across different communities.
+Besides, ward and ward number belong to the nature of the distribution of public wifi in Toronto itself, and do not take into account the influencing factors to be explored, but provide a new idea that can be connected with more social information associated with ward to explore more influential factors that can be studied.
 
-Since I only focused on the impact of building type and public wifi layout in Toronto, I kept only the building type, ward, and ward number columns during the data cleaning process in order to quickly extract available data.
-In addition, I carried out basic data cleaning on column values and deleted building data and NA information that did not contain wifi to improve readability, then summaried the building types and used the able() of knitr to create @tbl-building:
+As my research mainly concentrated on the influence of building type and the layout of public WiFi in Toronto, I did the data cleaning process by retaining only the columns related to building type, ward, and ward number to efficiently extract relevant data.
+Additionally, I performed basic data cleaning on column values, removing non-WiFi-related building data and NA information to enhance clarity.
+Subsequently, I knitr to generate the @tbl-wifi table for analysis:
+```{r}
+#| label: tbl-wifi
+#| tbl-cap: Sample of cleaned wifi data
+#| echo: false
+
+cleaned_wifi_data %>%
+  head(10) %>%
+  kable(
+    col.names = c("Building Type", "Ward", "Ward Number"),
+    digits = 1
+  )
+
+```
+@tbl-wifi shows reoughly the building types, ward and ward number of all buildings with public wifi in Toronto.
+To facilitate a clearer understanding of the distribution of public WiFi across different building types, I grouped the data by building type and summarized the number of free WiFi hotspots associated with each category showed in @tbl-building.
 
@@
 ```
+This process involved grouping the data by building type and calculating the total count of free WiFi instances within each category.
+By organizing the data in this manner, it aimed to provide an overview of the prevalence of public WiFi across various types of buildings, including arenas, libraries, community recreation centers (CRC), Outdoor Pool Building (OPB), Artificial Ice Rink Building (AIRB), and Washroom Building (WB).
 
-From @tbl-final, there are 198 free public WiFi location in the City of Toronto in total.
+From @tbl-building, we know that there are 198 free public WiFi location in the City of Toronto in total.
 Further breakdown shows that there are  100 free public WiFi location that are set in library, 74 free public WiFi location that are set in Community Recreation Centre(CRC), 27 free public WiFi location that are set in Arena, 17 free public WiFi location that are set in Outdoor Pool Building (OPB).
@@
 
-Library has the highest number of free public WiFi, which means that, in comparison to other types of buildings in Toronto, libraries offer the highest availability or coverage of free public WiFi services.
-This suggests that libraries prioritize or invest in providing accessible and free internet connectivity to the public within Toronto.
-Besides, Community Recreation Centre has the second most amount of free public WiFi.
-This information is indicative of the efforts or infrastructure investments made by Toronto goverment to offer accessible and free internet connectivity to the public.
-
 ## Second Dataset: 2021 Ward Profiles based on the 25-Ward model
 
-In addition to building types, I also studied the population and income of each ward in Toronto as another two variables to deeply understand the factors affecting the distribution of public wifi in Toronto.
+In addition to building types, I also studied the "population" and "income" of each ward in Toronto as another two variables to deeply understand the factors affecting the distribution of public wifi in Toronto.
+In general, demographic characteristics such as age, household size, and cultural background can influence the demand for and usage patterns of public WiFi, and more people always mean more demand for the Internet.
+In addition, areas with lower median incomes may face greater barriers to internet access due to affordability issues or lack of infrastructure investment.
+
 Therefore, the ward profile dataset [@toronto-ward-profiles] published by City Planning based on the 2021 Census data was also included in the analysis.
 This dataset contains demographic, social and economic information such as income and population for each ward, and last updated on January 3, 2024.
-
-Ward profile data is stored in an Excel file with multiple tabs, and the relevant data used for this analysis is included as a variable in the first TAB, the 2021 Census.
 Data included population, age, race, language and median household income.
@@
 This research is concerned with the impact of population and income on the distribution of free wifi, so I only choose to analyze these two variables as my research objects, and created statistic to visualize the data.
-we were able to get initial valid information but still need more clarity for further data analysis, so I established a new variable, "number of free wifi", to quantify the performance of free wifi in different building types.
-Then I used kable() from knitr to get @tbl-final:
+we were able to get initial valid information but still need more clarity for further data analysis, so I established a new variable, "number of free wifi", to quantify the performance of free wifi in different building types, and then knitr to get @tbl-final:
 
@@
 
-From @tbl-final, there are 25 wards in total in the City of Toronto, and the wards with the highest population are: Etobicoke-Lakeshore (Ward 3) at 139920, Spadina-Fort York (Ward 10) at 135400.
-The wards with the lowest population are: Scarborough North (Ward 23) at 94025, Don Valley East (Ward 16) at 94335.
-The wards with the highest household income are: Spadina-Fort York (Ward 10) at $80730$, Toronto Centre (Ward 13) at $68965$.
-The wards with the lowest household income are: Humber River-Black Creek (Ward 7) at $65,458$, York South-Weston (Ward 5) at $67,964$.
+From @tbl-final, there are 25 wards in total in the City of Toronto, and the wards with the highest population is Etobicoke-Lakeshore (Ward 3) at 139920, and the wards with the lowest population is Scarborough North (Ward 23) at 94025.
+The wards with the highest household income is Spadina-Fort York (Ward 10) at $80730$, and the wards with the lowest household income is Humber River-Black Creek (Ward 7) at $65,458$.
+The wards with the highest household income is Spadina-Fort York (Ward 10) at $80730$, and the wards with the lowest household income is Humber River-Black Creek (Ward 7) at $65,458$.
+The wards with the biggest number of Free public wifi is Humber River-Black Creek (Ward 7) at 14, and the wards with the smallest number of Free public wifi are Etobicoke Centre (Ward 2) at 5, Willowdale (Ward 18) at 5.
 
 It intuitively shows the social developments of 25 neighbourhoods in Toronto, including population, median income, and number of wifi, providing data support for the following study of the linear relationship between the three.
-It is important to note that, according to the database, more populated neighborhoods do not necessarily have more median income, which laid the foundation for our decision to separate the two variables to analyze the impact of public wifi distribution.
+It is important to note that, according to the database, more populated neighborhoods do not necessarily have more median income, as well as the number of free wifi.
+This observation underscores the complexity of urban dynamics and highlights the need to analyze each variable independently to understand its impact on public WiFi distribution.
 
@@
 ```
-@fig-interest shows the data visualization shows that the ward1, 3, and 7 have significantly more than the median number of public wifi, 12, 13, and 14 respectively, which is more than 1/3 times more than the average of 8.
-From this graph, we can observe that except for the ward 1, 3 and 7, the number of public wifi in other wards is basically in the same range, and there is no particularly serious shortage.
-This indicates that the level of public Internet accessibility strength is generally consistent across wards in Toronto.
+@fig-interest illustrates the disparities in public WiFi distribution across Toronto's wards, highlighting variations in coverage levels.
+Specifically, Ward 1, Ward 3, and Ward 7 stand out with substantially higher numbers of public WiFi hotspots compared to the median value of 8.
+Ward 1, with 12 hotspots, Ward 3, with 13 hotspots, and Ward 7, with 14 hotspots, each exceed the median by more than 1/3, indicating significantly more extensive coverage in these areas.
+In contrast, the number of public WiFi hotspots in other wards remains relatively consistent and does not exhibit any significant shortages.
+This observation suggests that while certain wards benefit from robust public WiFi infrastructure, others may experience disparities in access, highlighting the need for targeted interventions to promote equitable Interncet connectivity across all neighborhoods in Toronto.
+
+\newpage
 
@@
 
-The goal of my modelling strategy is twofold.
-Firstly, I want to analyze to what effect power play opportunities, power play effectiveness, save percentage, goals against average and shooting percentages all have on average goals scored per game.
-Secondly, I want to use these results to predict the average goals scored per game for the next season.
+After conducting data analysis of the two dataset, we have observed that there might be a correlation between the population of a ward and the number of free public WiFi in this ward, as well as the median income of a ward and the number of free public WiFi in this ward.
+To gain further insights and give resaonable advise about future Toronto public free wifi distribution, we construct two linear regression models, tested the two hypotheses separately.
 
-In the section below, the Bayesian analysis model used to investigate the multiple linear regression model of average goals scored per game is detailed.
+In the section below, the final models are detailed investigated here.
 
@@
 
-Define $y_i$ as the leaguewide average goals scored per game.
-Then $\beta_1$ is the leaguewide average power plays per game, $\gamma_1$ is leaguewide powerplay percentage, $\theta_1$ is leaguewide save percentage, $\phi_1$ is the leaguewide goals against average, and $\kappa_1$ is the leaguewide average shooting percentage.
+Define $Y$ as the dependent variable, which is the number of free public WiFi in a ward, and $X$ as the independent variable, which is the population of a ward in the first model, or the median income of a ward in the second model.
+Then $\beta_0$ is the intercept, representing the expected number of free WiFi hotspots when the population or income is zero, $\beta_1$ is the slope coefficient, indicating the change in the number of free WiFi hotspots for a one-unit increase in population or in income.
+$\epsilon$ is the error term, representing the random variability in the number of free WiFi hotspots that is not explained by population or income.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \theta_i + \phi_i + \kappa_i + \beta_i * \gamma_i \\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\theta &\sim \mbox{Normal}(0, 2.5) \\
-\phi &\sim \mbox{Normal}(0, 2.5) \\
-\kappa &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in `R` [@citeR] using the `rstanarm` package of [@citestan]. We use the default priors from `rstanarm`.
+$$ 
+Y =  \beta_{0} + \beta_{1}X + ε
+$$
 
@@
 
-We expect a positive relationship between average goals scored per game and power play opportunities, power play percentage, goals against average, and shooting percentage.
-The higher those four variables are, the more goals there should be.
-In particular, there should also be a negative relationship between goals scored and save percentage.
-With a lower save percentage, there are less saves that are occuring which means there should be more goals.
-Additionally, the two power play variables are interacted with eachother, since they are directly related to one another.
-There cannot be a power play percentage without a power play opportunity.
+Our goal of the linear regression model is to estimate the values of $\beta_0$ and $\beta_1$ that minimize the sum of squared errors between the number of  public free wifi of a ward ($Y$) and the population or the median income correspounding ward ($X$) in the data sets.
+By minimizing this sum of squared errors, the linear regression model can produce the best-fit line to describe the linear relationship between the two variables, allowing for the estimation of $Y$ values based on known $X$ values.
+
+In addition, the statistical significance of $\beta_1$  can be assessed using a t-test, which tests whether the estimated coefficient is significantly different from zero.
+If the p-value of the t-test is less than a chosen significance level, we can conclude that there is a significant relationship between the population of a ward and the number of free public WiFi in this ward, as well as the median income of a ward and the number of free public WiFi in this ward.
+
+\newpage
 
@@
 **Societal need for public Wi-Fi differs according to building types**
-
-While I talked about the increase of the amount of productive forwards and defensemen, here I will discuss the increase in the amount of good forwards and compare them to the amount of elite forwards.
-To be classified as an elite forward, usually they would need to finish the season at around a 1.4 PPG pace.
-Because of how hard of a feat it is to achieve a sustained 1.4 PPG pace, there are very few elite forwards in the NHL.
-@fig-building shows the amount of forwards that finish in multiple PPG ranges.
 
@@
 ```
-
-Library has the highest number of free public WiFi, which means that, in comparison to other types of buildings in Toronto, libraries offer the highest availability or coverage of free public WiFi services.
-This suggests that libraries prioritize or invest in providing accessible and free internet connectivity to the public within Toronto.
-Besides, Community Recreation Centre has the second most amount of free public WiFi.
-This information is indicative of the efforts or infrastructure investments made by Community Recreation Centres to offer accessible and free internet connectivity to the public.
+From @fig-building, we can find out significant trends regarding the distribution of free WiFi across different building types in Toronto.
+Libraries emerge as crucial hubs of connectivity, boasting the highest number of free WiFi hotspots with a total of 99 instances.
+This underscores the pivotal role that libraries play in providing accessible internet resources to the community, serving as essential spaces for education, research, and digital inclusion initiatives.
+Additionally, Community Recreation Centers (CRC) also exhibit a substantial presence of free WiFi, with 50 hotspots identified.
+These centers serve as vital community hubs where residents can engage in recreational activities, access social services, and participate in educational programs.
 
 The type of building can often reflect the different needs of modern people for WiFi, and having more public WiFi coverage proves that it has a stronger degree of Internet digitization needs.
-
 For example, libraries have embraced technology, expanding their offerings beyond books to include a rich array of digital resources.
@@
 Here, individuals can check emails, work remotely, and stay informed about current events, all while enjoying a cup of coffee.
+
+The prevalence of free WiFi in libraries and CRCs highlights their significance as inclusive spaces that support lifelong learning, social interaction, and access to digital resources for all members of the community.
+This underscores the importance of investing in these institutions to ensure equitable access to internet connectivity and promote digital inclusion in urban environments.
 
@@
 ```
-The first linear regression model, as illustrated in @fig-1, explores the relationship between the number of free public WiFi locations and the population of wards across Toronto.
+The first linear regression model, as illustrated in @fig-1, explores the relationship between the number of free public WiFi locations in a ward and the population of corresponding ward across Toronto, which is a slight positive correlation.
 The scatterplot depicts the general trend of increased number of public wifi as bigger population, further emphasized by the regression line.
@@
 
-However, the coefficient for population (0.00) indicates that there is no statistically significant linear relationship between the population of a ward and the number of free public WiFi locations it hosts, as the coefficient's p-value exceeds conventional thresholds for significance.
+@tbl-model1 displays the coefficients of the predictor variables of the linear regression model.
+The coefficient for population (0.00) from @tbl-model1 indicates that there is no statistically significant linear relationship between the population of a ward and the number of free public WiFi locations it hosts, as the coefficient's p-value exceeds conventional thresholds for significance.
 This finding suggests that the population size alone may not be a strong predictor of the availability of free public WiFi in Toronto's wards.
@@
 
-In summary, while the model attempts to assess the relationship between population and the availability of free public WiFi, its findings indicate that population alone is not a strong predictor of the number of free public WiFi locations in Toronto's wards.
-
-From @fig-1, we can clearly see that Etobicoke-Lakeshore (Ward 3) has the greatest population, and Spadina-Fort York (Ward 10) has the second greatest population.
+In summary, while we attempts to assess the relationship between population and the availability of free public WiFi, its findings indicate that population alone is not a strong predictor of the number of free public WiFi locations in Toronto's wards, but there is still a slightly positive correlation between the two.
 
@@
 ```
-
-@tbl-model1 displays the coefficients of the predictor variables of the linear regression model.
-All the variables included in the model have p-values less than 0.05, indicating that they have a significant impact on the permanent immigrant inflows rate.
-Therefore, all the variables are retained in the final model.
-For every 1% increase in the Employment Rate, there is a 0.04% increase in the Permanent Inflows Rate.
-The relationship between Employment Rate and Permanent Inflows Rate is positive, which means that as the Employment Rate increases, the Permanent Inflows Rate also tends to increase.
-The slope coefficient of 0.04 indicates that for every 1% increase in the Employment Rate, we can expect a 0.04% increase in the Permanent Inflows Rate, all else being equal.
-As the "Year" variable increases by 1, the estimated permanent immigrant inflow increases by 0.04% on average, holding other predictor variables constant.
-However, it is important to note that this is a correlation and does not necessarily imply causation.
-Other factors, such as government policies and global economic conditions, may also have an impact on Permanent Inflows Rate.
-
-In fact, this is similar to the distribution of free public WiFi in Toronto, but it does not exactly match.
-Ward 3, as the most populous area, has 13 public WiFi, which is a relatively large number in all Wards, and Ward 10 has 5, which is an insignificant number.
-But overall, for different Wards, a higher population slightly results in a higher amount of public WiFi.
 
@@
   labs(
-    x = "Average Income of a Ward",
+    x = "Median Income of a Ward",
     y = "Number of Free Wifi in a ward"
@@
 ```
-From @fig-2, we can clearly see that Spadina-Fort York (Ward 10) has the greatest average income, Toronto Centre (Ward 13) has the second greatest average income, and Etobicoke-Lakeshore (Ward 3) has the third greatest average income.
+The first linear regression model, as illustrated in @fig-2, explores the relationship between the number of free public WiFi locations in a ward and the median income of corresponding ward in Toronto, which shows very limited correlation.
+The scatterplot shows that all the points distributed very scattered and disordered, and cannot find a general trend of increased or decreased number of public wifi as bigger population.
+Also the regression line is almost parallel to the x-axis.
+The model indicates that there is nearly no relationship between the median income of a ward and the number of free public WiFi locations it hosts.
+
+@tbl-model2 displays the coefficients of the predictor variables of the linear regression model.
+The coefficient for income is 0.00, suggesting that there is virtually no change in the number of free WiFi locations with an increase in median income.
+Additionally, the R-squared value of 0.001 indicates that only a negligible proportion of the variance in the number of free WiFi locations can be explained by median income.
+The adjusted R-squared value of -0.043 further suggests that the model's predictive power is very low.
+
+Overall, these findings suggest that median income alone may not be a significant predictor of the number of free WiFi locations in a ward, highlighting the need for further exploration of other factors that may influence public WiFi distribution in urban areas.
 
@@
 modelsummary::modelsummary(
-  list("Average Income of a Ward" = income_model),
+  list("Median Income of a Ward" = income_model),
   fmt = 2
@@
 
-@tbl-model2 displays the coefficients of the predictor variables of the linear regression model.
-All the variables included in the model have p-values less than 0.05, indicating that they have a significant impact on the permanent immigrant inflows rate.
-Therefore, all the variables are retained in the final model.
-For every 1% increase in the Employment Rate, there is a 0.04% increase in the Permanent Inflows Rate.
-The relationship between Employment Rate and Permanent Inflows Rate is positive, which means that as the Employment Rate increases, the Permanent Inflows Rate also tends to increase.
-The slope coefficient of 0.04 indicates that for every 1% increase in the Employment Rate, we can expect a 0.04% increase in the Permanent Inflows Rate, all else being equal.
-As the "Year" variable increases by 1, the estimated permanent immigrant inflow increases by 0.04% on average, holding other predictor variables constant.
-However, it is important to note that this is a correlation and does not necessarily imply causation.
-Other factors, such as government policies and global economic conditions, may also have an impact on Permanent Inflows Rate.
+\newpage
 
-We can conclude that there is no observable correlation between the number of Toronto public WiFi and the average household income in a ward.
-
-\newpage
+Based one the two linear regression models, we can conclude that the population of a ward has slightly positive correlation with the number of Toronto public wifi, but there is no observable correlation between the number of Toronto public wifi and the median household income in a ward.
 
@@
 
+## Findings and advice
+
+The findings of this study shed light on the intricate relationship between public WiFi distribution and urban connectivity in Toronto.
+Understanding these findings is crucial for policymakers, urban planners, and community advocates seeking to bridge the digital divide and foster a more equitable society.
+
+**Equitable Access to Public WiFi: A Necessity, Not a Luxury**
+
+One of the central themes emerging from our analysis is the importance of equitable access to public WiFi.
+While public WiFi hotspots are widely available across Toronto, our study reveals disparities in their distribution.
+Certain building types, such as libraries, community centers, and parks, emerge as hubs of connectivity, while others lag behind.
+This discrepancy underscores the need for deliberate efforts to ensure that all neighborhoods, regardless of socioeconomic status, have access to reliable internet connectivity.
+
+**Addressing Differences caused by population density**
+
+Another significant finding is the variation in public WiFi distribution across population density found by linear regression model.
+Although median income of a ward doesn't play a big role in the distribution of public free wifi, wards with bigger population tend to have more extensive coverage compared to less crowed places.
+This disparity suggests that access to public WiFi is not evenly distributed across the city, potentially exacerbating existing inequalities, since fewer public wifi leads to less Internet conections and usages, making it increasingly difficult for networks development in the region.
+To address this issue, policymakers should prioritize investment in underserved communities, ensuring that all residents have access to essential digital resources.
+
+**Implications for Urban Policy**
+
+The findings of this study have several implications for urban policy and planning.
+Firstly, policymakers should prioritize the equitable distribution of public WiFi infrastructure, particularly in underserved neighborhoods.
+This may involve targeted investment in community centers, libraries, and public spaces to ensure that all residents have access to reliable internet connectivity.
+
+Secondly, urban planners should integrate considerations of digital equity into their decision-making processes.
+By prioritizing projects that promote connectivity and digital inclusion, cities can create more resilient and equitable communities.
+
+Finally, community engagement and collaboration are essential for addressing the digital divide.
+By involving local residents in the planning and implementation of public WiFi initiatives, cities can ensure that solutions are tailored to the specific needs of each community.
+
+## Bias and ethical concerns
+
+\newpage
+## Weaknesses & Next Steps
+
+\newpage
 All in all, a significant number of people are increasingly relying on public Wi-Fi when they're out and about, underscoring their need for instant, hassle-free connectivity.
@@
 
+In conclusion, the findings of this study underscore the importance of equitable access to public WiFi in fostering urban connectivity and social inclusion.
+By addressing disparities in WiFi distribution and prioritizing digital equity in urban policy and planning initiatives, cities can create more resilient, connected, and equitable communities for all residents.
+
+Also, by targeting building types with limited connectivity, policymakers can work towards improving the wifi  coverage in relatively backward buildings and promoting inclusivity in urban development efforts.
+can work towards closing the digital divide and promoting inclusivity in urban development efforts.
+
 # Appendix 

### 2024-04-18 outputs/paper/paper.qmd
improve the writing of discussion part

@@
 
-The findings of this study shed light on the intricate relationship between public WiFi distribution and urban connectivity in Toronto.
-Understanding these findings is crucial for policymakers, urban planners, and community advocates seeking to bridge the digital divide and foster a more equitable society.
+The findings of this study shed light on the intricate relationship between public WiFi distribution and urban social characteristics in Toronto.
+The intricate dynamics surrounding public WiFi distribution in Toronto encompass a multitude of factors that interact to shape urban wareless Internet connectivity.
+Beyond simplistic correlations, this study delves into the interplay between the distribution of public wifi, income, population, and building types.
+
+Firstly, while median income is often presumed to be a determining factor in access to resources, this study challenges that assumption.
+The negligible impact of income on the availability of free WiFi locations suggests that other factors may hold greater significance.
+This could include the allocation of resources by municipal authorities, the presence of community initiatives, or the density of public amenities, all of which may influence WiFi accessibility independently of income levels.
+
+Secondly, the study sheds light on the role of infrastructure and community engagement in shaping public WiFi distribution.
+For instance, areas with higher population densities or greater concentrations of public amenities, such as libraries or community centers, may naturally attract more public WiFi installations due to higher foot traffic and demand.
+Additionally, community-led initiatives or partnerships between local organizations and internet service providers could contribute to the expansion of WiFi networks in underserved areas.
+
+Moreover, the study underscores the importance of considering the diverse needs and preferences of communities when planning WiFi infrastructure.
+Factors such as cultural diversity, language barriers, and digital literacy levels may influence the effectiveness of public WiFi initiatives and require tailored solutions to ensure inclusivity.
+
+Overall, the intricate dynamics surrounding public WiFi distribution in Toronto highlight the need for a comprehensive and context-sensitive approach to urban connectivity.
+By understanding and addressing the multifaceted factors at play, policymakers, urban planners, and community advocates can develop more effective strategies to bridge the digital divide and foster a truly equitable society.Understanding these findings is crucial for policymakers, urban planners, and community advocates seeking to bridge the digital divide and foster a more equitable society.
 
@@
 
+As we delve into the analysis of public WiFi distribution in Toronto, it is important to navigate through various biases and ethical considerations that may influence our findings and subsequent policy recommendations.
+We explore three points that underscore the importance of addressing bias and ethical concerns in our research.
+
+**Measurement Bias**
+
+An important consideration is measurement bias, which arises from inaccuracies or inconsistencies in measuring or reporting variables such as population, median income, and the number of free WiFi locations.
+For example, relying on self-reported revenue data can introduce errors and discrepancies across wards.
+To address measurement bias, we must employ reliable data collection methods and validate our findings by cross-referencing with multiple sources.
+
+**Sampling Bias**
+
+If the sample size is not representative of the Toronto population, or if certain groups are excluded or underrepresented in our analysis, then sampling bias poses a significant risk to the validity of our findings.
+By ensuring that our sample is inclusive and representative, we can mitigate the effects of sampling bias and draw more accurate conclusions about the relationship between population, income, and public wifi distribution.
+
+**Ethical Considerations**
+
+Ethical considerations are crucial in any research effort, especially when it comes to the collection and analysis of personal data.
+In the context of public WiFi distribution, we must prioritize the privacy of individuals, including anonymizing personal information, and adhere to strict privacy regulations to safeguard the confidentiality and rights of participants.
+
+In conclusion, navigating bias and ethical considerations is essential for producing meaningful and actionable insights into research of public WiFi distribution in Toronto.
+By acknowledging these challenges, we can enhance the validity and integrity of our research findings and contribute to the development of more equitable policies.
+
 \newpage
@@
 
-\newpage
-All in all, a significant number of people are increasingly relying on public Wi-Fi when they're out and about, underscoring their need for instant, hassle-free connectivity.
-Whether for work, social interaction or navigation, the modern lifestyle requires a seamless Internet experience, and the increasing number of academic and recreational projects joining the ranks of the application of public free WiFi is undoubtedly a strong support for the development of public WiFi in Toronto.
+A significant limitation of our study is the complexity of the factors that influence the distribution of public WiFi in Toronto.
+While we strive to gather comprehensive data sets on population, median income, and the number of free WiFi locations, it is important to recognize that the influencing factors are multifaceted and interdependent.
+The true determinants of public WiFi distribution likely include a wide range of socioeconomic, infrastructural, and community-related factors that are difficult to accurately capture through quantitative analysis alone.
 
-Furthermore, our data analysis indicates that while population size does exert a modest impact on the dissemination of free WiFi, the average income of a region lacks sufficient influence to be regarded as a determining factor in the allocation of public WiFi.
-
-In conclusion, the findings of this study underscore the importance of equitable access to public WiFi in fostering urban connectivity and social inclusion.
-By addressing disparities in WiFi distribution and prioritizing digital equity in urban policy and planning initiatives, cities can create more resilient, connected, and equitable communities for all residents.
-
-Also, by targeting building types with limited connectivity, policymakers can work towards improving the wifi  coverage in relatively backward buildings and promoting inclusivity in urban development efforts.
-can work towards closing the digital divide and promoting inclusivity in urban development efforts.
+To address this limitation, we must employ a holistic approach that combines qualitative research methods, stakeholder interviews, and community engagement initiatives to gain a deeper understanding of the potential drivers of public WiFi distribution.
+Through a combination of quantitative and qualitative analysis, we can spot subtle dynamic changes and develop more comprehensive strategies to improve public WiFi connectivity in Toronto.
 

### 2024-04-18 outputs/paper/paper.qmd
add enhancement for appendix

@@
 # Appendix
+
+For enhancement, I take advantage of shiny, which is to quickly add some interactivity to our graphs according to the simulated data.
+Here(@fig-shiny) is the screnshot of my shiny work, and the interactive graphis is in the folder "Enhancement - Shiny Plot".
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-shiny
+knitr::include_graphics("Shiny.png")
+```
+
 \newpage

### 2024-04-18 Enhancement - Shiny Plot/app.R

@@ -1,0 +1,62 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    https://shiny.posit.co/
+#
+
+library(shiny)
+library(tidyverse)
+
+simulated_data <-
+  read_csv(
+    file = here("inputs/data/simulated_data.csv"),
+    show_col_types = FALSE
+  )
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+
+    # Application title
+    titlePanel("Building with Wifi Data"),
+
+    # Sidebar with a slider input for number of bins 
+    sidebarLayout(
+        sidebarPanel(
+            sliderInput("bins",
+                        "Number of bins:",
+                        min = 1,
+                        max = 50,
+                        value = 30)
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+           plotOutput("distPlot")
+        )
+    )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+
+    output$distPlot <- renderPlot({
+        # generate bins based on input$bins from ui.R
+      simulated_data |> ggplot(aes(x= building_number, fill= has_wifi))+
+        geom_histogram(position = "dodge", bins = input$bins)+
+        theme_minimal()+
+        scale_fill_brewer(palette = "Set1") +
+        labs(
+          x = "Building Number",
+          y = "Occurrences",
+          fill = "has wifi"
+        )
+        # draw the histogram with the specified number of bins
+       
+    })
+}
+
+# Run the application 
+shinyApp(ui = ui, server = server)

### 2024-04-18 scripts/00-simulate_data.R

@@ -19,8 +19,8 @@ library(tidyverse)
 set.seed(85)
 
 simulated_data <-
-  tibble(
-    has_wifi = rep(x = "Y", times = 100),
+  tibble(building_number = c(1:100),
+    has_wifi = sample(c("Y","N"), size = 100, replace = TRUE, prob = c(0.9, 0.1)),
     # Based on: https://www.toronto.ca/city-government/accountability-operations-customer-service/long-term-vision-plans-and-strategies/smart-cityto/internet-connectivity-connectto/connectto-free-public-wifi-locations/
     building_type = c(
       rep(x = "library", times = 20),

### 2024-04-18 scripts/00-simulate_data.R
add simulate data

@@ -36,3 +36,24 @@ head(simulated_data)
 #### Save data ####
 write_csv(simulated_data,"inputs/data/simulated_data.csv")
 
+#### Simulate final data ####
+set.seed(85)
+
+simulated_data <-
+  tibble(building_number = c(1:100),
+         has_wifi = sample(c("Y","N"), size = 100, replace = TRUE, prob = c(0.9, 0.1)),
+         # Based on: https://www.toronto.ca/city-government/accountability-operations-customer-service/long-term-vision-plans-and-strategies/smart-cityto/internet-connectivity-connectto/connectto-free-public-wifi-locations/
+         building_type = c(
+           rep(x = "library", times = 20),
+           rep(x = "Arena", times = 40),
+           rep(x = "Community", times = 40)
+         ),
+         ward = rep(c(x = "Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"), times = 4),
+         ward_number =rep(c(1:25), times = 4)
+  )
+
+head(simulated_data)
+
+#### Save final data ####
+write_csv(simulated_data,"inputs/data/simulated_data.csv")
+

### 2024-04-18 outputs/paper/paper.qmd
correct a typo
Delete unnecessary model

@@
 # IMPORT MODEL FROM OUTPUTS FOLDER
-wifi_model <- readRDS(here::here("outputs/models/wifi_model.rds"))
 population_model <- readRDS(here::here("outputs/models/population_model.rds"))
@@
 
-The findings of this study shed light on the intricate relationship between public WiFi distribution and urban social characteristics in Toronto.
+The findings of this study shed light on the intricate relationship between public Wifi distribution and urban social characteristics in Toronto.
 The intricate dynamics surrounding public WiFi distribution in Toronto encompass a multitude of factors that interact to shape urban wareless Internet connectivity.
@@
 
+\newpage
 # Appendix
@@
 For enhancement, I take advantage of shiny, which is to quickly add some interactivity to our graphs according to the simulated data.
-Here(@fig-shiny) is the screnshot of my shiny work, and the interactive graphis is in the folder "Enhancement - Shiny Plot".
+Here(@fig-shiny) is the screnshot of my shiny work, and the interactive graph is is in the folder "Enhancement - Shiny Plot".
 

### 2024-04-18 scripts/04-model.R

@@ -14,24 +14,11 @@ library(modelsummary)
 #### Read data ####
 final_data <- read.csv(here::here("outputs/data/final_data.csv"))
 
-#### Model data 1 ####
-wifi_model <-
-  stan_glm(
-    formula = Number_of_free_wifi ~ Population + Income,
-    data = final_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 123
-  )
-
-
 ## creating simulated data for the model
 set.seed(853)
 
 
-#### Linear Model data 2 ####
+#### Linear Model data 1 ####
 population_model <-
   lm(
     Number_of_free_wifi ~ Population,
@@ -42,7 +29,7 @@ summary(population_model)
 modelsummary(population_model)
 
 
-#### Linear Model data 3 ####
+#### Linear Model data 2 ####
 income_model <-
   lm(
     Number_of_free_wifi ~ Income,
@@ -54,18 +41,13 @@ modelsummary(income_model)
 
 #### Save model ####
 
-saveRDS(
-  wifi_model,
-  file = here::here("outputs/models/wifi_model.rds")
-)
-
-# LINEAR MODEL 2
+# LINEAR MODEL 1
 saveRDS(
   population_model,
   file = "outputs/models/population_model.rds"
 )
 
-# LINEAR MODEL 3
+# LINEAR MODEL 2
 saveRDS(
   income_model,
   file = "outputs/models/income_model.rds"

### 2024-04-18 outputs/paper/paper.qmd
do some small corrections

@@
 ---
-title: "Urban Connectivity Dynamics: Analyzing Free Public WiFi Distribution in Toronto"
-subtitle: "By Building Typology, Population, and Socioeconomic Indicators"
+title: "Urban Digital Development: A Study of Public Free WiFi Distribution in Toronto"
+subtitle: "Insights from Building Type, Population and Income"
 author: 
@@
 date-format: long
-abstract: "Public WiFi is a vital tool in modern society, offering no-cost internet access and supporting economic, social and educational development.
-In response to the increasing importance of public free wifi in contemporary urban life, this paper investigates the distribution and influencing factors of free public WiFi in Toronto.
-Drawing on datasets from Open Data Toronto, the paper focuses on the association between the building types of the place with public wifi, population and income of each ward in Toronto and number of  Toronto public free wifi by creating linear regression models, aimed to shed light on the correlation between free public WiFi distribution and building types, population, and income, and put forward relevent suggestions for the future development.
-Understanding these findings is crucial for policymakers, urban planners, and community advocates developing digital public Internet and foster a more equitable society."
+abstract: "Public WiFi is a vital tool for modern society, supporting economic, social, and educational development through free internet access.
+This study examines the influencing factors of free public WiFi distribution in Toronto using datasets from Open Data Toronto.
+Specifically, it investigates the relationship between building types, population, median income of each ward in Toronto, and the number of public free WiFi hotspots.
+Linear regression models are utilized to analyze these correlations and offer insights into the current distribution of public WiFi, including potential benefits and disadvantages.
+Findings indicate that areas with higher population densities or greater concentrations of public amenities tend to have more public WiFi installations, while median income does not significantly influence public WiFi deployment.
+Understanding these findings is crucial for policymakers and urban planners to ensure the equitable development of digital public Internet infrastructure."
 format: pdf
@@
 
-In an era characterized by an ever-deepening reliance on digital technologies, the intricate tapestry of urban internet connectivity is emerging as a pivotal force in the transformation of our societal fabric.
-Countries globally are acknowledging internet access as an indispensable tool for active participation in contemporary democratic societies.
-In Canada, the Canadian Radio-television and Telecommunications Commission (CRTC) has recognized that "broadband internet access services are crucial to Canada’s economic, social, democratic, and cultural fabric."(@citiesfordigitalrights-toronto-wifi) As a university student navigating the complexities of contemporary living, the significance of understanding these dynamics cannot be overstated.
-We are acutely aware of the transformative potential that connectivity holds.
+In the digital age, access to the internet has become increasingly essential for participation in economic, social, and educational activities.
+With the continuous progress of science and technology, public WiFi, which can be easily and wirelessly connected to the Internet at a low cost, is gradually improving and promoting to all corners of the world.
+Nowadays, public free WiFi has emerged as a crucial tool in bridging the digital divide and promoting equitable access to online resources.
+It fosters internet accessibility, enabling people to stay connected without constraints.
+It also improve the access to information, job searching, and remote work opportunities are further facilitated, enhancing the overall quality of life.
+In emergencies, public WiFi could also serve as a crucial communication lifeline, demonstrating its indispensable role in connecting people and providing access to vital services.(@broadbandsearch-public-wifi)  In Canada, the Canadian Radio-television and Telecommunications Commission (CRTC) has recognized that "broadband internet access services are crucial to Canada’s economic, social, democratic, and cultural fabric."(@citiesfordigitalrights-toronto-wifi)
 
-Public WiFi, often taken for granted in our daily lives, emerges as a vital tool shaping the contours of a technologically inclusive society.
-It plays a pivotal role in modern society, offering a myriad of benefits to individuals.
-Beyond providing a cost-effective alternative to cellular data, it fosters internet accessibility, enabling people to stay connected without constraints.
-This flexibility and mobility are particularly advantageous for students, travelers, and those on a budget, allowing them to work, study, or communicate from diverse locations.
-Moreover, public WiFi contributes to social inclusion, bridging the digital divide and ensuring that even those without home internet access can engage in online activities.
-Access to information, job searching, and remote work opportunities are further facilitated, enhancing the overall quality of life.
-In emergencies, public WiFi serves as a crucial communication lifeline, demonstrating its indispensable role in connecting people and providing access to vital services.(@broadbandsearch-public-wifi)
+Recognizing its significance, this paper embarks on an exploration of the distribution of free public WiFi in Toronto, and delves into its potential influencing factors, aiming to shed light on the dynamics shaping its accessibility across different urban areas.
+By examining the relationship between building types, population demographics, median income, and the number of public free WiFi hotspots in different wards of Toronto, this study seeks to uncover the underlying factors driving the spatial distribution of public WiFi infrastructure.
+Utilizing datasets sourced from Open Data Toronto, our analysis employs rigorous statistical methods, including linear regression models, to explore the correlations between various socio-economic variables and the availability of public WiFi.
+Through this approach, we aim to offer insights into the benefits and potential drawbacks of current public WiFi distribution strategies, with a particular focus on understanding the role of population density and the presence of public amenities in driving WiFi deployment.
+By elucidating these factors, we aim to provide valuable analysis for policymakers and urban planners tasked with ensuring the equitable provision of digital public Internet infrastructure.
 
-Beyond the conveniences of internet access, it serves as a linchpin fostering economic development, facilitating education, and fortifying social connections.
-The present study, anchored in the evolving landscape of Toronto, endeavors to unravel the mysteries surrounding the distribution of public WiFi.
-This paper embarks on a compelling exploration of the distribution of free public WiFi in Toronto, recognizing it as not merely a technological amenity but a cornerstone of our modern existence.
-By examining the impact of various factors, such as building types, demographic patterns, and economic influences, this research seeks to paint a comprehensive portrait of the role and impact of public WiFi in our lives, providing readers with a lens to comprehend the symbiotic relationship between technology and urban living.
+Our findings reveal intriguing patterns regarding the relationship between population density, public amenity concentration, and the presence of public WiFi installations.
+Notably, areas with higher population densities or greater concentrations of public amenities tend to exhibit a higher prevalence of public WiFi hotspots.
+However, our analysis also suggests that median income levels do not significantly influence the deployment of public WiFi infrastructure.
+These findings underscore the importance of considering diverse socio-economic factors in shaping strategies for the equitable development of digital public Internet infrastructure in urban areas.
+This paper contributes to our understanding of the factors influencing the distribution of public WiFi in urban settings.
+By elucidating the relationship between socio-economic variables and public WiFi deployment, our study offers valuable insights that can inform the development of more equitable and inclusive digital infrastructure policies, and also provide readers with a lens to comprehend the symbiotic relationship between technology and urban living.
 
-This research is divided into three main parts: `Introduction`(@sec-intro), `Data`(@sec-data), `Model` (@sec-model), `Results` (@sec-results), and `Discussion` (@sec-disc).
-Commencing with the Data section, a meticulous exposition ensues, delineating the origin of datasets garnered from the OpenDataToronto Library and expounding upon the procedural rigors applied for data refinement and analysis.
-The conclusion part shows what I found during the analysis, while goes deeper into those findings, and finally wraps up the main discoveries from this paper about public free WiFi in Toronto.
+This research is divided into five main parts: `Introduction`(@sec-intro), `Data`(@sec-data), `Model` (@sec-model), `Results` (@sec-results), and `Discussion` (@sec-disc).
+@sec-data provides a detailed explanation of the two databases used for analysis, data processing, including visualization of variables of interest.
+@sec-model proposes a simple linear regression model to be applied to our two variables of population and income separately.
+@sec-results presents the findings from the analysis, and @sec-disc takes a deeper look at those findings, concluding with a summary of all of the analysis in this paper on the distribution of public free WiFi in Toronto and a discussion of bias and next steps.
 
@@
 
-In this section, I'll explore the R-based data collection process and generate the visualizations in the RStudio environment [@citeR;@citeRStudio].
+In this section, We explore the R-based data collection process and generate the visualizations in the RStudio environment [@citeR;@citeRStudio].
 All the datasets studied in this paper come from "Open Data Toronto" [@citeODT], a data source allowing researchers and policymakers to access and utilize the data which can be downloaded for free from its website and GitHub repository, in order to finish various statistical analyses, research studies, and decision-making processes.
-The two main datasets used in this paper are @toronto_wifi_dataset, which provides descriptive data such as the location and building type of Toronto free public wifi, and @toronto_ward_profiles, which contains economic, population and other social information of all 25 wards in Toronto collected in 2021.
+The two main datasets used in this paper are @toronto_wifi_dataset, which provides descriptive data such as the location and building type of Toronto free public WiFi, and @toronto_ward_profiles, which contain economic, population and other social information of all 25 wards in Toronto collected in 2021.
 
@@
 
-@toronto_wifi_dataset, the dataset I primarily use to generate visualizations is an open source tool designed to be publicly available and to encourage the development of valuable insights.
-It published by Toronto's Department of Information Technology, collects location information and descriptive data for free public WiFi locations in the City of Toronto and is updated monthly, with the latest update on April 15, 2024.
-These data include address, building type, building name, public name, zip code, whether there is public wifi, owner, ward, ward number, and other geographical information.
+The dataset @toronto_wifi_dataset, is an open-source resource provided by Toronto's Department of Information Technology.
+It collects location information and descriptive data for free public WiFi locations in the City of Toronto and is regularly updated, with the latest update on April 15, 2024.
+This dataset includes various details such as address, building type, building name, zip code, and ward information.
 
-Among all statistical data about free public wifi, our analysis is particularly interested in "building type", "ward" and "ward number", which are considered as the main subject variables of our study.
-Analyzing public WiFi distribution by building type allows us to understand where internet access is concentrated and which locations may be underserved.
-This insight can inform decisions about where to expand infrastructure, and how to ensure equitable access to digital resources across different communities.
-Besides, ward and ward number belong to the nature of the distribution of public wifi in Toronto itself, and do not take into account the influencing factors to be explored, but provide a new idea that can be connected with more social information associated with ward to explore more influential factors that can be studied.
+Of particular interest in our analysis are the variables "building type," "ward," and "ward number," which serve as the main focus of our study.
+Analyzing public WiFi distribution by building type helps us identify areas with concentrated internet access and those that may require additional infrastructure development to ensure equitable access across communities.
+Similarly, examining ward and ward number distributions provides insights into the natural distribution of public WiFi in Toronto, laying the groundwork for further exploration of associated social factors.
 
-As my research mainly concentrated on the influence of building type and the layout of public WiFi in Toronto, I did the data cleaning process by retaining only the columns related to building type, ward, and ward number to efficiently extract relevant data.
-Additionally, I performed basic data cleaning on column values, removing non-WiFi-related building data and NA information to enhance clarity.
-Subsequently, I knitr to generate the @tbl-wifi table for analysis:
+To streamline our analysis, we conducted data cleaning by retaining only relevant columns related to building type, ward, and ward number.
+Additionally, we performed basic data cleaning procedures to remove non-WiFi-related building data and NA values, enhancing the clarity and accuracy of our analysis.
+Subsequently, we utilized knitr to generate the @tbl-wifi table for further analysis:
 ```{r}
@@
 ```
-@tbl-wifi shows reoughly the building types, ward and ward number of all buildings with public wifi in Toronto.
-To facilitate a clearer understanding of the distribution of public WiFi across different building types, I grouped the data by building type and summarized the number of free WiFi hotspots associated with each category showed in @tbl-building.
+@tbl-wifi shows reoughly the building types, ward and ward number of all buildings with public WiFi in Toronto.
+To facilitate a clearer understanding of the distribution of public WiFi across different building types, we grouped the data by building type and summarized the number of free WiFi hotspots associated with each category showed in @tbl-building.
 
@@
 Further breakdown shows that there are  100 free public WiFi location that are set in library, 74 free public WiFi location that are set in Community Recreation Centre(CRC), 27 free public WiFi location that are set in Arena, 17 free public WiFi location that are set in Outdoor Pool Building (OPB).
-Since the same wifi may cover multiple buildings, the sum of the number of public wifi corresponding to each building type is greater than the total number of public wifi in Toronto.
-The four building types mentioned above have the largest number of public wifi, covering almost 90% of total.
+Since the same WiFi may cover multiple buildings, the sum of the number of public WiFi corresponding to each building type is greater than the total number of public WiFi in Toronto.
+The four building types mentioned above have the largest number of public WiFi, covering almost 90% of total.
+\newpage
 
@@
 
-In addition to building types, I also studied the "population" and "income" of each ward in Toronto as another two variables to deeply understand the factors affecting the distribution of public wifi in Toronto.
-In general, demographic characteristics such as age, household size, and cultural background can influence the demand for and usage patterns of public WiFi, and more people always mean more demand for the Internet.
-In addition, areas with lower median incomes may face greater barriers to internet access due to affordability issues or lack of infrastructure investment.
+We expanded our analysis beyond building types to include population and income data for each ward in Toronto.
+Demographic factors such as age, household size, and cultural background can influence the demand for and usage patterns of public WiFi, with higher population densities generally indicating greater internet demand.
+Additionally, areas with lower median incomes may face barriers to internet access due to affordability issues or insufficient infrastructure investment.
 
-Therefore, the ward profile dataset [@toronto-ward-profiles] published by City Planning based on the 2021 Census data was also included in the analysis.
-This dataset contains demographic, social and economic information such as income and population for each ward, and last updated on January 3, 2024.
-Data included population, age, race, language and median household income.
-Median household income is the pre-tax income a family receives from employment, investment and government support.
-Median income is used because it is less sensitive to very high and very low income values than the average.
+To incorporate this data, we utilized the ward profile dataset [@toronto-ward-profiles] published by City Planning, based on 2021 Census data and last updated on January 3, 2024.
+This dataset includes demographic, social, and economic information such as population, age, race, language, and median household income.
+Median household income, representing pre-tax income a family receives from employment, investment and government support.
+Median income is used in our research because it is less sensitive to very high and very low income values than the average.
 
-This research is concerned with the impact of population and income on the distribution of free wifi, so I only choose to analyze these two variables as my research objects, and created statistic to visualize the data.
-we were able to get initial valid information but still need more clarity for further data analysis, so I established a new variable, "number of free wifi", to quantify the performance of free wifi in different building types, and then knitr to get @tbl-final:
-
+Focusing on the impact of population and income on free WiFi distribution, we narrowed our analysis to these two variables.
+While initial data provided valuable insights, we sought further clarity through digital analysis.
+To quantify the performance of free WiFi across different building types, we created a new variable by combining it with the first dataset, labeled "number of free WiFi," then getting @tbl-final to facilitate deeper investigation:
 ```{r}
@@
 
-From @tbl-final, there are 25 wards in total in the City of Toronto, and the wards with the highest population is Etobicoke-Lakeshore (Ward 3) at 139920, and the wards with the lowest population is Scarborough North (Ward 23) at 94025.
-The wards with the highest household income is Spadina-Fort York (Ward 10) at $80730$, and the wards with the lowest household income is Humber River-Black Creek (Ward 7) at $65,458$.
-The wards with the highest household income is Spadina-Fort York (Ward 10) at $80730$, and the wards with the lowest household income is Humber River-Black Creek (Ward 7) at $65,458$.
-The wards with the biggest number of Free public wifi is Humber River-Black Creek (Ward 7) at 14, and the wards with the smallest number of Free public wifi are Etobicoke Centre (Ward 2) at 5, Willowdale (Ward 18) at 5.
+@tbl-final reveals key insights into the population, median income and amount of public WiFi of Toronto's 25 wards.
+It highlights variations in population size, median income, and the number of free WiFi hotspots across these wards.
+For instance, Etobicoke-Lakeshore (Ward 3) boasts the highest population count at 139,920, while Scarborough North (Ward 23) records the lowest at 94,025.
+Similarly, Spadina-Fort York (Ward 10) ranks highest in median household income at 80,730 dollars, contrasting with Humber River-Black Creek (Ward 7), which reports the lowest income at 65,458 dollars.
+In terms of free WiFi availability, Humber River-Black Creek (Ward 7) leads with 14 hotspots, whereas both Etobicoke Centre (Ward 2) and Willowdale (Ward 18) have the fewest, each with 5.
 
-It intuitively shows the social developments of 25 neighbourhoods in Toronto, including population, median income, and number of wifi, providing data support for the following study of the linear relationship between the three.
-It is important to note that, according to the database, more populated neighborhoods do not necessarily have more median income, as well as the number of free wifi.
+It intuitively shows the social developments of 25 neighbourhoods in Toronto, including population, median income, and number of WiFi, providing data support for the following study of the linear relationship between the three.
+It is important to note that, according to the database, more populated neighborhoods do not necessarily have more median income, as well as the number of free WiFi.
 This observation underscores the complexity of urban dynamics and highlights the need to analyze each variable independently to understand its impact on public WiFi distribution.
@@
 #| label: fig-interest
-#| fig-cap: The number of free Wifi available in different wards in Toronto
+#| fig-cap: The number of free WiFi available in different wards in Toronto
 #| echo: false
@@
   labs(
-    title = "The number of free Wifi available in different wards in Toronto",
+    title = "The number of free WiFi available in different wards in Toronto",
     x = "Ward Number",
-    y = "Number of Free Public Wifi"
+    y = "Number of Free Public WiFi"
   ) 
@@
 After conducting data analysis of the two dataset, we have observed that there might be a correlation between the population of a ward and the number of free public WiFi in this ward, as well as the median income of a ward and the number of free public WiFi in this ward.
-To gain further insights and give resaonable advise about future Toronto public free wifi distribution, we construct two linear regression models, tested the two hypotheses separately.
+To gain further insights and give resaonable advise about future Toronto public free WiFi distribution, we construct two linear regression models, tested the two hypotheses separately.
 
@@
 
-Our goal of the linear regression model is to estimate the values of $\beta_0$ and $\beta_1$ that minimize the sum of squared errors between the number of  public free wifi of a ward ($Y$) and the population or the median income correspounding ward ($X$) in the data sets.
+Our goal of the linear regression model is to estimate the values of $\beta_0$ and $\beta_1$ that minimize the sum of squared errors between the number of  public free WiFi of a ward ($Y$) and the population or the median income correspounding ward ($X$) in the data sets.
 By minimizing this sum of squared errors, the linear regression model can produce the best-fit line to describe the linear relationship between the two variables, allowing for the estimation of $Y$ values based on known $X$ values.
@@
 #| label: fig-building
-#| fig-cap: The number of free Wifi in Toronto by building type
+#| fig-cap: The number of free WiFi in Toronto by building type
 #| echo: false
@@
     x = "Building Types",
-    y = "Number of Free Public Wifi"
+    y = "Number of Free Public WiFi"
   )
@@
 #| label: fig-1
-#| fig-cap: Linear regression model of the number of Free public wifi by the populations of all wards in Toronto.
+#| fig-cap: Linear regression model of the number of Free public WiFi by the populations of all wards in Toronto.
 #| echo: false
@@
     x = "Population of a Ward",
-    y = "Number of Free Wifi in a ward"
+    y = "Number of Free WiFi in a ward"
   ) +
@@
 The first linear regression model, as illustrated in @fig-1, explores the relationship between the number of free public WiFi locations in a ward and the population of corresponding ward across Toronto, which is a slight positive correlation.
-The scatterplot depicts the general trend of increased number of public wifi as bigger population, further emphasized by the regression line.
-The model indicates that as the population of a ward gets bigger, the number of free public wifi increases as well.
+The scatterplot depicts the general trend of increased number of public WiFi as bigger population, further emphasized by the regression line.
+The model indicates that as the population of a ward gets bigger, the number of free public WiFi increases as well.
 
@@
 #| label: fig-2
-#| fig-cap: Linear regression model of the number of Free public wifi by the Average Income of all wards in Toronto.
+#| fig-cap: Linear regression model of the number of Free public WiFi by the Average Income of all wards in Toronto.
 #| echo: false
@@
     x = "Median Income of a Ward",
-    y = "Number of Free Wifi in a ward"
+    y = "Number of Free WiFi in a ward"
   ) +
@@
 ```
-The first linear regression model, as illustrated in @fig-2, explores the relationship between the number of free public WiFi locations in a ward and the median income of corresponding ward in Toronto, which shows very limited correlation.
-The scatterplot shows that all the points distributed very scattered and disordered, and cannot find a general trend of increased or decreased number of public wifi as bigger population.
+The second linear regression model, as illustrated in @fig-2, explores the relationship between the number of free public WiFi locations in a ward and the median income of corresponding ward in Toronto, which shows very limited correlation.
+The scatterplot shows that all the points distributed very scattered and disordered, and cannot find a general trend of increased or decreased number of public WiFi as bigger population.
 Also the regression line is almost parallel to the x-axis.
@@
 
-Based one the two linear regression models, we can conclude that the population of a ward has slightly positive correlation with the number of Toronto public wifi, but there is no observable correlation between the number of Toronto public wifi and the median household income in a ward.
+Based one the two linear regression models, we can conclude that the population of a ward has slightly positive correlation with the number of Toronto public WiFi, but there is no observable correlation between the number of Toronto public WiFi and the median household income in a ward.
+
+\newpage
 
@@
 The intricate dynamics surrounding public WiFi distribution in Toronto encompass a multitude of factors that interact to shape urban wareless Internet connectivity.
-Beyond simplistic correlations, this study delves into the interplay between the distribution of public wifi, income, population, and building types.
+Beyond simplistic correlations, this study delves into the interplay between the distribution of public WiFi, income, population, and building types.
 
@@
 
-Overall, the intricate dynamics surrounding public WiFi distribution in Toronto highlight the need for a comprehensive and context-sensitive approach to urban connectivity.
-By understanding and addressing the multifaceted factors at play, policymakers, urban planners, and community advocates can develop more effective strategies to bridge the digital divide and foster a truly equitable society.Understanding these findings is crucial for policymakers, urban planners, and community advocates seeking to bridge the digital divide and foster a more equitable society.
+Overall, the intricate dynamics surrounding public WiFi distribution in Toronto highlight the need for a and context-sensitive approach to urban connectivity.
+By understanding and addressing the multifaceted factors at play, policymakers, urban planners, and community advocates can develop more effective strategies to bridge the digital divide and foster a truly equitable society.
+
+Here are some extra analysis as advice:
 
@@
 Another significant finding is the variation in public WiFi distribution across population density found by linear regression model.
-Although median income of a ward doesn't play a big role in the distribution of public free wifi, wards with bigger population tend to have more extensive coverage compared to less crowed places.
-This disparity suggests that access to public WiFi is not evenly distributed across the city, potentially exacerbating existing inequalities, since fewer public wifi leads to less Internet conections and usages, making it increasingly difficult for networks development in the region.
+Although median income of a ward doesn't play a big role in the distribution of public free WiFi, wards with bigger population tend to have more extensive coverage compared to less crowed places.
+This disparity suggests that access to public WiFi is not evenly distributed across the city, potentially exacerbating existing inequalities, since fewer public WiFi leads to less Internet conections and usages, making it increasingly difficult for networks development in the region.
 To address this issue, policymakers should prioritize investment in underserved communities, ensuring that all residents have access to essential digital resources.
@@
 This may involve targeted investment in community centers, libraries, and public spaces to ensure that all residents have access to reliable internet connectivity.
-
 Secondly, urban planners should integrate considerations of digital equity into their decision-making processes.
 By prioritizing projects that promote connectivity and digital inclusion, cities can create more resilient and equitable communities.
-
 Finally, community engagement and collaboration are essential for addressing the digital divide.
@@
 If the sample size is not representative of the Toronto population, or if certain groups are excluded or underrepresented in our analysis, then sampling bias poses a significant risk to the validity of our findings.
-By ensuring that our sample is inclusive and representative, we can mitigate the effects of sampling bias and draw more accurate conclusions about the relationship between population, income, and public wifi distribution.
+By ensuring that our sample is inclusive and representative, we can mitigate the effects of sampling bias and draw more accurate conclusions about the relationship between population, income, and public WiFi distribution.
 
@@
 
-\newpage
 ## Weaknesses & Next Steps
@@
 A significant limitation of our study is the complexity of the factors that influence the distribution of public WiFi in Toronto.
-While we strive to gather comprehensive data sets on population, median income, and the number of free WiFi locations, it is important to recognize that the influencing factors are multifaceted and interdependent.
+While we strive to gather data sets on population, median income, and the number of free WiFi locations, it is important to recognize that the influencing factors are multifaceted and interdependent.
 The true determinants of public WiFi distribution likely include a wide range of socioeconomic, infrastructural, and community-related factors that are difficult to accurately capture through quantitative analysis alone.
@@
 To address this limitation, we must employ a holistic approach that combines qualitative research methods, stakeholder interviews, and community engagement initiatives to gain a deeper understanding of the potential drivers of public WiFi distribution.
-Through a combination of quantitative and qualitative analysis, we can spot subtle dynamic changes and develop more comprehensive strategies to improve public WiFi connectivity in Toronto.
+Through a combination of quantitative and qualitative analysis, we can spot subtle dynamic changes and develop more strategies to improve public WiFi connectivity in Toronto.
 

### 2024-04-18 scripts/00-simulate_data.R
Add a new simulated dataset

@@ -18,6 +18,7 @@ library(tidyverse)
 #### Simulate data ####
 set.seed(85)
 
+#This simulation data mainly simulates the preliminary data obtained through simple cleaning
 simulated_data <-
   tibble(building_number = c(1:100),
     has_wifi = sample(c("Y","N"), size = 100, replace = TRUE, prob = c(0.9, 0.1)),
@@ -39,21 +40,16 @@ write_csv(simulated_data,"inputs/data/simulated_data.csv")
 #### Simulate final data ####
 set.seed(85)
 
-simulated_data <-
-  tibble(building_number = c(1:100),
-         has_wifi = sample(c("Y","N"), size = 100, replace = TRUE, prob = c(0.9, 0.1)),
-         # Based on: https://www.toronto.ca/city-government/accountability-operations-customer-service/long-term-vision-plans-and-strategies/smart-cityto/internet-connectivity-connectto/connectto-free-public-wifi-locations/
-         building_type = c(
-           rep(x = "library", times = 20),
-           rep(x = "Arena", times = 40),
-           rep(x = "Community", times = 40)
-         ),
-         ward = rep(c(x = "Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"), times = 4),
-         ward_number =rep(c(1:25), times = 4)
+# The simulation of final data mainly simulates the efficient data obtained after combining two databases
+simulated_final_data <-
+  tibble(Population = rpois(n = 25, lambda = 110000),
+         Income = rpois(n = 25, lambda = 48000),
+         WardNumber = c(1:25),
+         Number_of_free_wifi = rpois(n = 25, lambda = 10)
   )
 
-head(simulated_data)
+head(simulated_final_data)
 
 #### Save final data ####
-write_csv(simulated_data,"inputs/data/simulated_data.csv")
+write_csv(simulated_final_data,"inputs/data/simulated_final_data.csv")
 

### 2024-04-18 scripts/03-test_data.R
Add more tests to examine the simulated final data

@@ -9,14 +9,97 @@
 #### Workspace setup ####
 library(tidyverse)
 
-#### Test data ####
+#### Test simulated final data ####
 
-data = read_csv("outputs/data/simulated_data.csv")
+data = read_csv("inputs/data/simulated_final_data.csv")
 
-# Test1
-all(data$ward_number >= 1 & data$ward_number <= 25)
+# Test 1: Check if Population values are non-negative
+if (all(simulated_final_data$Population >= 0)) {
+  print("All population values are non-negative.")
+} else {
+  print("Some population values are negative!")
+}
 
-# Test2
-all(data$ward %in% c("Etobicoke North", "Etobicoke Centre", "Etobicoke-Lakeshore", "Parkdale-High Park", "York South-Weston", "York Centre", "Humber River-Black Creek", "Eglinton-Lawrence", "Davenport", "Spadina-Fort York", "University-Rosedale", "Toronto-St. Paul's", "Toronto Centre", "Toronto-Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches-East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough-Agincourt", "Scarborough North", "Scarborough-Guildwood", "Scarborough-Rouge Park"))
+# Test 2: Check if Median Income values are within a reasonable range
+if (all(simulated_final_data$Income >= 0)) {
+  print("All median income values are non-negative.")
+} else {
+  print("Some median income values are negative!")
+}
 
-# Test3
+# Test 3: Check if Number of Free Wi-Fi spots are non-negative
+if (all(simulated_final_data$Number_of_free_wifi >= 0)) {
+  print("All number of free Wi-Fi values are non-negative.")
+} else {
+  print("Some number of free Wi-Fi values are negative!")
+}
+
+# Test 4: Check if all Population values are integers
+if (all.equal(simulated_final_data$Population, as.integer(simulated_final_data$Population))) {
+  print("All population values are integers.")
+} else {
+  print("Some population values are not integers!")
+}
+
+# Test 5: Check if Median Income values are within a reasonable range (e.g., not exceeding $200,000)
+if (all(simulated_final_data$Income <= 200000)) {
+  print("All median income values are within a reasonable range.")
+} else {
+  print("Some median income values exceed the reasonable range!")
+}
+
+# Test 6: Check if Ward Numbers are unique
+if (length(unique(simulated_final_data$WardNumber)) == nrow(simulated_final_data)) {
[…]
